### 2014-04-07 final_presentation/presentation.tex
Added blueprint files for presentation and paper.

\documentclass[10pt]{beamer}

\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{attrib}
\usepackage{textpos}

\usetheme{Dresden}
\usecolortheme{dove}
\usenavigationsymbolstemplate{}
\title{Research?}
\author{Pieter Brederode, Timo Koppenberg, \\ Jelco Bodewes, Anton Golov}
\institute{B3OMI}

\begin{document}

\begin{frame}
    \maketitle
\end{frame}

\section{Introduction}
\begin{frame}{Overview}
    \begin{itemize}
        \item We did research!
        \item That's it, really.
    \end{itemize}
\end{frame}

\section{Conclusions}

\begin{frame}{Results}
    \begin{itemize}
        \item What results?
        \item What are you talking about?
    \end{itemize}
\end{frame}

\end{document}

### 2014-04-07 paper/paper.tex
Added blueprint files for presentation and paper.

\documentclass[12pt,a4paper]{article}
% vim: set textwidth=100:

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{hyperref}

\title{Data Structures Comparative Review Thing}
\author{Pieter Brederode, Timo Koppenberg, \\ Jelco Bodewes, Anton Golov}

\begin{document}
    \maketitle

    \begin{abstract}
        We did research and got results.
    \end{abstract}

    \section{Introduction}

    Ph'nglui mglw'nafh Cthulhu R'lyeh wgah'nagl fhtagn~\cite{CallOfCthulhu}. Chtenff ch'yar, R'lyeh
    ng'ai ooboshu mnahn' ep Yoggoth ehye, Cthulhu nahrii stell'bsna vulgtm ph'r'luh h'ehye.
    Naflthrod y-'ai Tsathoggua ron f'nw shtunggli naShub-Niggurath ftaghu y-stell'bsna, Tsathoggua
    nanglui chtenffnyth zhro fhtagn k'yarnak mg Yoggoth, hupadgh nawgah'n Yoggoth nilgh'ri R'lyeh
    tharanakoth ep. 'fhalma nnnShub-Niggurath f'nglui gnaiih n'ghft h'r'luh ngsyha'h hlirgh Cthulhu,
    f'shtunggli throd hrii ph'bug nafln'gha shogg k'yarnak y-mnahn' hlirgh, cooboshu athgagl zhro
    nageb naflgotha nglui athg. Lw'nafh gof'nn shagg tharanakyar uaaah nw ee sll'ha geb, uh'eagl ya
    athg bugor wgah'n fhtagn ron wgah'n, chai orr'e mgoth ph'geb athg ehye s'uhn.

    \section{Results}

    Ilyaa lw'nafh f'sll'ha mg hupadgh li'heeog n'gha athg lloignyth geb, Dagonnyth 'bthnk r'luh
    sll'ha naflgrah'n gnaiih Shub-Niggurath Tsathoggua stell'bsna f'tharanak, sll'ha nw ep
    ngstell'bsna f'zhro naflgotha y'hah 'ai. Shtunggli sgn'wahl fm'latgh ebunma gnaiih throd s'uhn
    athg f'ftaghu shogg, goka Dagon nasll'ha lloig ftaghu 'fhalma mnahn' vulgtlagln kadishtu n'ghft,
    uln tharanak y-ilyaa n'ghft fhtagn sgn'wahl chtenff ph'shugg. Ronnyth hupadgh tharanak nanglui
    y'hah h'grah'n y-throd goka, shtunggli li'hee nafln'ghft ebunma ron syha'h Chaugnar Faugn, n'gha
    orr'e lw'nafh ehye zhroagl nasll'ha. Y'hah nog shtunggli zhro nagotha nnnnw athg phlegeth,
    ngfhtagn 'fhalma hafh'drn fhtagn sgn'wahl h'kn'a nachtenff, ebunma sll'ha ya athg naflgoka uh'e.

    K'yarnak kadishtu shogg li'heeyar fm'latgh zhro uh'e phlegeth 'bthnk goka fhtagn geb Azathoth
    geb, Shub-Niggurath ftaghu zhro shogg vulgtlagln ya nnnchtenff gof'nn orr'e nnnshtunggli ron.
    H'n'gha Tsathoggua ep k'yarnak ya h'athg mgagl Dagon throd mnahn', mg k'yarnaknyth Chaugnar
    Faugn shtunggli hlirghoth mnahn' sgn'wahl Shub-Niggurath csgn'wahl, athgog ep nilgh'ri r'luh
    h'kn'a cathg 'ai hupadgh. Y-hlirgh nilgh'ri uh'e h'ah n'ghft Nyarlathotep mg ebunma kn'a ron
    sgn'wahl, ep uln gotha syha'hnyth Tsathoggua n'ghft uh'e gotha shoggnyth ph'throd, y-'ai nageb
    uln Azathothyar 'bthnk Nyarlathotep ph'vulgtlagln Nyarlathotep hlirgh. Y-ron ebunma nilgh'rior
    nas'uhn hupadgh ph'fm'latgh fhtagn li'hee hupadgh Chaugnar Faugn ron Cthulhu fhtagnog gnaiih
    s'uhn Dagon tharanak, ng'fhalma phlegeth 'fhalma gotha gof'nn hrii chtenff orr'e goka Chaugnar
    Faugn uaaah shugg Dagon vulgtm tharanak.

    Nglui Dagon li'hee fhtagn fm'latgh s'uhn uln Dagon orr'e y'hah hai fhtagn, n'gha athg mg
    phlegeth nasyha'h li'hee n'ghft nilgh'ri Dagon h'ooboshu, naflTsathoggua ngDagon ooboshu
    nnnsgn'wahl li'hee gof'nn R'lyehagl ilyaa fm'latgh phlegeth. Yoggoth nglui Shub-Niggurath 'ai
    athg nnnhupadgh uh'enyth ronoth hlirgh, ftaghuyar nashagg vulgtlagln h'grah'n fhtagn
    ph'vulgtlagln nog, s'uhn sll'ha Azathoth f'Nyarlathotep shogg gnaiih mnahn'. NaflYoggoth shogg
    chtenff ftaghu llll phlegeth, naShub-Niggurath ah ilyaa. Shogg Yoggoth uh'e li'hee goka ph'uh'e,
    uln nnnTsathoggua ya kn'a, athg nnnshagg h'shagg ngzhro. Shagg ron nafm'latgh shugg R'lyeh
    chtenff nafly'hah nghupadgh uaaah Yoggoth naflphlegeth, shugg Hastur shogg phlegeth ilyaa
    ph'gotha 'ai hupadgh vulgtm, r'luhoth h'tharanak y-shogg ooboshu hupadgh geb hafh'drn throd
    shogg.

    \section{Conclusions}

    Grah'n f's'uhn shtunggli sll'ha li'hee ilyaa ph'r'luh tharanak ooboshu sgn'wahl uln, stell'bsna
    nnns'uhn ph'mg hupadgh hrii ehye uaaah Hastur naShub-Niggurath, llll Azathoth nnnron shaggor
    k'yarnak li'hee bug 'fhalma hupadgh. Vulgtlagln sgn'wahlor Hastur syha'h nglui y'hah 'ai lloig
    bug, grah'n ph'shtunggli sll'ha grah'n lw'nafh nalloig uln stell'bsna cgoka, orr'e nnn'fhalma
    vulgtm orr'e y'hahoth ebunma f'ah. Ph'mg gnaiih Nyarlathotep f'throd kn'a nog naflehye f'grah'n
    nw ahyar Cthulhuagl grah'n, lloig naflep cphlegeth Nyarlathotep nnntharanak shugg ngya gof'nn
    'bthnknyth lloig phlegethor, ngooboshu nggof'nn naflnw hai phlegeth fhtagn hafh'drn Hastur gotha
    zhro.

    \bibliographystyle{alpha}

    \bibliography{paper}

\end{document}

### 2014-04-07 data.cpp
Made output more SPSS-friendly.

@@ -4,10 +4,15 @@
 
 std::ostream& operator<<(std::ostream& os, Result const& result) {
     print_aligned(result.structure, os);
+    os << ',';
     print_aligned(result.key_type, os);
+    os << ',';
     print_aligned(result.elements, os);
+    os << ',';
     print_aligned(result.insertion_time.count(), os);
+    os << ',';
     print_aligned(result.query_time.count(), os);
+    os << ',';
     print_aligned(result.memory_usage, os);
     return os;
 }

### 2014-04-07 main.cpp

@@ -67,12 +67,12 @@ int main(int argc, char* argv[]) {
     auto data_domains = read_lines(std::ifstream(argv[1]));
     auto data_fullpaths = read_lines(std::ifstream(argv[2]));
 
-    print_aligned("structure");
-    print_aligned("key");
-    print_aligned("entries");
-    print_aligned("insert");
-    print_aligned("query");
-    print_aligned("memory");
+    print_aligned("MapType");
+    print_aligned("KeyType");
+    print_aligned("Entries");
+    print_aligned("Insert");
+    print_aligned("Query");
+    print_aligned("Memory");
     std::cout << std::endl;
 
     std::ostream_iterator<Result> output(std::cout, "\n");

### 2014-04-09 main.cpp
Minor cleanup.

@@ -68,10 +68,15 @@ int main(int argc, char* argv[]) {
     auto data_fullpaths = read_lines(std::ifstream(argv[2]));
 
     print_aligned("MapType");
+    std::cout << ',';
     print_aligned("KeyType");
+    std::cout << ',';
     print_aligned("Entries");
+    std::cout << ',';
     print_aligned("Insert");
+    std::cout << ',';
     print_aligned("Query");
+    std::cout << ',';
     print_aligned("Memory");
     std::cout << std::endl;
 

### 2014-04-09 measure.hpp

@@ -1,7 +1,6 @@
 #pragma once
 
 #include "data.hpp"
-#include "misc.hpp"
 #include "measuring_allocator.hpp"
 
 #include <algorithm>

### 2014-04-09 misc.cpp

@@ -1,5 +1,0 @@
-#include "misc.hpp"
-
-long get_memory_usage() {
-    return 0;
-}

### 2014-04-09 misc.hpp

@@ -1,3 +1,0 @@
-#pragma once
-
-long get_memory_usage();

### 2014-04-09 rng/CSRNG.cpp

@@ -1,27 +1,0 @@
-/*
-   Description: Static variables for CS::RNG
-
-   Copyright 2010 Cleo Saulnier
-
-   Licensed under the Apache License, Version 2.0 (the "License");
-   you may not use this file except in compliance with the License.
-   You may obtain a copy of the License at
-
-       http://www.apache.org/licenses/LICENSE-2.0
-
-   Unless required by applicable law or agreed to in writing, software
-   distributed under the License is distributed on an "AS IS" BASIS,
-   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
-   See the License for the specific language governing permissions and
-   limitations under the License.
-*/
-
-#include "CSRNG.h"
-
-namespace CS
-{
-
-r250* RNG::rng = NULL;
-int RNG::counter = 0;
-
-}

### 2014-04-09 rng/CSRNG.h

@@ -1,77 +1,0 @@
-/*
-   Description: Header for CS::RNG
-
-   Copyright 2010 Cleo Saulnier
-
-   Licensed under the Apache License, Version 2.0 (the "License");
-   you may not use this file except in compliance with the License.
-   You may obtain a copy of the License at
-
-       http://www.apache.org/licenses/LICENSE-2.0
-
-   Unless required by applicable law or agreed to in writing, software
-   distributed under the License is distributed on an "AS IS" BASIS,
-   WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
-   See the License for the specific language governing permissions and
-   limitations under the License.
-*/
-#ifndef CSRNGH
-#define CSRNGH
-
-#include <stdio.h>
-#include "r250.h"
-
-namespace CS
-{
-
-class RNG
-{
-private:
-  static r250 *rng;
-  static int counter;
-public:
-  RNG()
-  {
-    counter++;
-    if (rng==NULL)
-    {
-      rng=new r250();
-      rng->init();
-    }
-  }
-
-  ~RNG()
-  {
-    counter--;
-    if (counter==0)
-    {
-      delete rng;
-      rng = NULL;      
-    }
-  }
-
-  unsigned int rand() {return rng->rand();}
-  double drand() {return rng->drand();}
-};
-
-class RNG_threaded
-{
-private:
-  r250 rng;
-public:
-  RNG_threaded()
-  {
-    rng.init();
-  }
-
-  ~RNG_threaded()
-  {
-  }
-
-  unsigned int rand() {return rng.rand();}
-  double drand() {return rng.drand();}
-};
-
-}
-
-#endif

### 2014-04-09 rng/moa.cpp

@@ -1,58 +1,0 @@
-/***************************************************************************
-                          moa.cpp  -  description
-                             -------------------
-    begin                : Thu Mar 7 2002
-    copyright            : (C) 2002-2010 by Cleo Saulnier
-    desc                 : MOA RNG algorithm
-
-    Notes:
-    Class to be used as the array initialiser for the r250 algorithm.
-
-    Algorithm created by Dr. George Marsaglia.
-
-    Dec 2010:
-    I remember writing the code based on the equations.
-    So I'm leaving my contribution in the public domain.
-
- ***************************************************************************/
-
-//#include "pch.h"
-
-#include "moa.h"
-//---------------------------------------------------------------------------
-
-moa_rng::moa_rng()
-{
-}
-
-moa_rng::~moa_rng()
-{
-}
-
-void moa_rng::init(unsigned int seed)
-{
-  for (int i=0;i<4;i++)
-  {
-    seed ^= seed << 13; seed ^= seed >> 17; seed ^= seed << 5;
-    X[i] = seed;
-  }
-  seed ^= seed << 13; seed ^= seed >> 17; seed ^= seed << 5;
-  C = seed;
-}
-
-unsigned int moa_rng::rand()
-{
-  unsigned long long int S = (unsigned long long int)2111111111 * (unsigned long long int)X[0];
-  S += (unsigned long long int)1492 * (unsigned long long int)X[1];
-  S += (unsigned long long int)1776 * (unsigned long long int)X[2];
-  S += (unsigned long long int)5115 * (unsigned long long int)X[3];
-  S += (unsigned long long int)C;
-  X[0] = X[1];
-  X[1] = X[2];
-  X[2] = X[3];
-  X[3] = (unsigned int)S;
-  C = (unsigned int)(S>>32);
-  return X[3];
-}
-
-

### 2014-04-09 rng/moa.h

@@ -1,35 +1,0 @@
-/***************************************************************************
-                          moa.h  -  description
-                             -------------------
-    begin                : Thu Mar 7 2002
-    copyright            : (C) 2002-2010 by Cleo Saulnier
-    desc                 : MOA RNG algorithm
-
-    Notes:
-    Class to be used as the array initialiser for the r250 algorithm.
-
-    Algorithm created by Dr. George Marsaglia.
-
-    Dec 2010:
-    I remember writing the code based on the equations.
-    So I'm leaving my contribution in the public domain.
-    
- ***************************************************************************/
-#ifndef moaH
-#define moaH
-/*! \brief MOA RNG Algorithm.
- *
- */
-class moa_rng
-{
-private:
-  unsigned int X[4];
-  unsigned int C;
-public:
-  moa_rng();
-  ~moa_rng();
-  void init(unsigned int seed);
-  unsigned int rand();
-};
-#endif
-

### 2014-04-09 rng/r250.cpp

@@ -1,133 +1,0 @@
-/***************************************************************************
-                          r250.cpp  -  description
-                             -------------------
-
-    begin                : ~2001
-    copyright            : (C) 2001-2010 by Cleo Saulnier
-    desc                 : r250 RNG class.
-
-    Notes:
-    Class to support r250 RNG algorithm.
-    Produces full 32 bit random numbers or normalized double value.
-
-    Code is from Maier in Dr. Dobbs Journal, 1991.
-    I adapted it for 32 bit and ported to C++ class.
-    I'm leaving my contribution in the public domain.
-
-    Here is the comment from the original code.
-***************************************************************************
-* Module: r250.cpp Description: implements R250 random number generator,
-* from S. Kirkpatrick and E. Stoll, Journal of Computational Physics, 40,
-* p. 517 (1981). Written by: W. L. Maier
-***************************************************************************
-
-    Maier was contacted and said,
-    "Yes, you can use the r250 code however you like."
-
-    Usage:
-    1. Create instance.  ex: r250 *r = new r250();  *or* r250 r;
-    2. Initialize seed.  ex: r->init(time(NULL));
-    3. Get random values.  ex: int number = r->rand();  *or* double number = r->drand();
-    4. delete instance if dynamic allocation.  ex: delete r;
-
- ***************************************************************************/
-//#include "pch.h"
-#include <stdio.h>
-#include <time.h>
-
-#include "r250.h"
-#include "moa.h"
-//---------------------------------------------------------------------------
-
-r250::r250()
-{
-}
-
-r250::~r250()
-{
-}
-
-// Auto init.
-// Uses clock as seed if available, time(NULL) otherwise.
-void r250::init()
-{
-  clock_t c = clock();
-  time_t t = time(NULL);
-  if (c==-1) init((unsigned int)t);
-  else init((unsigned int)(c+t));
-}
-
-void r250::init(unsigned int seed)
-{
-  int j, k;
-  unsigned long int mask, msb;
-  unsigned int x;
-
-	moa_rng *moa = new moa_rng();
-	moa->init(seed);
-
-	r250_index = 0;
-	for (j = 0; j < R250_SIZE; j++) // fill r250 buffer with 32 bit values
-	{
-		x = moa->rand();
-		r250_buffer[j] = x;
-	}
-	delete moa;
-	moa = NULL;
[…]

### 2014-04-09 rng/r250.h

@@ -1,63 +1,0 @@
-/***************************************************************************
-                           r250.h  -  description
-                             -------------------
-    begin                : ~2001
-    copyright            : (C) 2001-2010 by Cleo Saulnier
-    desc                 : r250 RNG class.
-
-    Notes:
-    Class to support r250 RNG algorithm.
-    Produces full 32 bit random numbers or normalized double value.
-
-    Code is from Maier in Dr. Dobbs Journal, 1991.
-    I adapted it for 32 bit and ported to C++ class.
-    I'm leaving my contribution in the public domain.
-
-    Here is the comment from the original code.
-***************************************************************************
-* Module: r250.cpp Description: implements R250 random number generator,
-* from S. Kirkpatrick and E. Stoll, Journal of Computational Physics, 40,
-* p. 517 (1981). Written by: W. L. Maier
-***************************************************************************
-
-    Maier was contacted and said,
-    "Yes, you can use the r250 code however you like."
-
-    Usage:
-    1. Create instance.  ex: r250 *r = new r250();  *or* r250 r;
-    2. Initialize seed.  ex: r->init(time(NULL));
-    3. Get random values.  ex: int number = r->rand();  *or* double number = r->drand();
-    4. delete instance if dynamic allocation.  ex: delete r;
-
- ***************************************************************************/
-#ifndef r250H
-#define r250H
-
-// These are default values for R250
-//#define R250_SIZE 250
-//#define R250_X2 103
-
-// For bigger cycles, use size=1279 and x2=418
-#define R250_SIZE 1279
-#define R250_X2 418
-
-/*! \brief R250 Algorithm.
- *
- */
-class r250
-{
-private:
-  unsigned int r250_buffer[R250_SIZE];
-  int r250_index;
-public:
-  r250();
-  ~r250();
-  void init();
-  void init(unsigned int seed);
-  unsigned int rand();
-  double drand();
-};
-
-
-#endif
-

### 2014-04-09 main.cpp
Fixed hashmap weird measurement issue.

@@ -82,13 +82,19 @@ int main(int argc, char* argv[]) {
 
     std::ostream_iterator<Result> output(std::cout, "\n");
 
+#ifndef NO_SKIPLIST
     eval_structure<Map<CS::KeyedSkipList, int>::type>("skiplist", "ip", data_ips, output);
     eval_structure<Map<CS::KeyedSkipList, std::string>::type>("skiplist", "domain", data_domains, output);
     eval_structure<Map<CS::KeyedSkipList, std::string>::type>("skiplist", "full_path", data_fullpaths, output);
+#endif
+#ifndef NO_BST
     eval_structure<Map<std::map, int>::type>("bst", "ip", data_ips, output);
     eval_structure<Map<std::map, std::string>::type>("bst", "domain", data_domains, output);
     eval_structure<Map<std::map, std::string>::type>("bst", "full_path", data_fullpaths, output);
+#endif
+#ifndef NO_HASHMAP
     eval_structure<Map<std::unordered_map, int>::type>("hashmap", "ip", data_ips, output);
     eval_structure<Map<std::unordered_map, std::string>::type>("hashmap", "domain", data_domains, output);
     eval_structure<Map<std::unordered_map, std::string>::type>("hashmap", "full_path", data_fullpaths, output);
+#endif
 }

### 2014-04-09 measure.hpp

@@ -78,14 +78,14 @@ template<template<template<typename> class> class MapTmpl, typename T, typename 
 std::vector<Result> eval_structure(std::string structure_name, std::string key_type, std::vector<T> const& data, OutIter out) {
     using clock = std::chrono::high_resolution_clock;
     std::vector<Result> results;
-    for (std::size_t i = 8; i < 21; ++i) {
+    for (std::size_t i = 8; i < 14; ++i) {
         auto const subset = random_subset(data, 1 << i);
         std::size_t const elements = 1 << i;
         using Map = MapTmpl<std::allocator>;
         using MemMeasureMap = MapTmpl<MeasuringAllocator>;
+        auto const memory_usage = eval_map_memory_usage<MemMeasureMap, clock>(subset.begin(), subset.end());
         auto const insert_result = eval_map_insertion<Map, clock>(subset.begin(), subset.end());
         auto const query_result = eval_map_query<Map, clock>(subset.begin(), subset.end());
-        auto const memory_usage = eval_map_memory_usage<MemMeasureMap, clock>(subset.begin(), subset.end());
         *out = {structure_name, key_type, elements, insert_result, query_result, memory_usage};
         ++out;
     }

### 2014-04-09 final_presentation/presentation.tex
Work on second part of final presentation

@@
 
-\section{Conclusions}
+\section{Results}
 
@@
     \begin{itemize}
-        \item What results?
-        \item What are you talking about?
+        \item 117 different setups.
+        \item 19 runs, for a total of 2223 cases.
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Amount of entries}
+	\begin{figure}
+	  \centering
+	    \includegraphics[width=0.95\textwidth]{SkiplistQuery}
+	  \caption{The average amount of time for each query operation, plotted against the total number of entries.}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Amount of entries}
+    \begin{itemize}
+        \item There is a positive correlation between the amount of entries and the average speed of all structures.
+        \item This is true for both insert and query average speed.
+        \item Not always true for average memory usage.
+    \end{itemize}       
+\end{frame}
+
+\begin{frame}{Amount of entries}
+    \begin{itemize}
+        \item Average memory usage for BST is constant.
+        \item Average memory usage for Hashmap has a positive correlation with the amount of entries.
+        \item Average memory usage for Skiplist has a negative correlation with the amount of entries.
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Amount of entries}
+	\begin{figure}
+	  \centering
+	    \includegraphics[width=0.95\textwidth]{SkiplistMemory}
+	  \caption{The average amount of time for each query operation, plotted against the total number of entries.}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Binary search tree}
+    \begin{itemize}
+        \item For Ip-Adresses: Significantly faster insertions than both other structures for low amount of entries.
+        \item Scales badly compared to the other structures.
+        \item Other keys: Always slower than Hashmap.
+    \end{itemize}       
+\end{frame}
+
+\begin{frame}{Hashmap}
+    \begin{itemize}
+        \item Significantly faster query speed in all situations compared to other structures.
+        \item Faster insert speed as well in most situations.
+        \item Highest amount of memory usage in all cases.
+    \end{itemize}       
+\end{frame}
+
+\begin{frame}{Skiplist}
+    \begin{itemize}
+        \item In no tested situations the fasted structure.
+        \item Similar performance to BST for large amount of entries.
+        \item Always the lowest memory usage, average memory usage goes down with more entries.
+    \end{itemize}       
+\end{frame}
+
+\section{Conclusions}
+
+\begin{frame}{Conclusions}
+    \begin{itemize}
+        \item Hashmap usually the best option.
+        \item Binary search tree better for small amounts of simple keys.
+        \item Skiplist useful for low memory usage.
     \end{itemize}

### 2014-04-09 final_presentation/presentation.tex
Wrote for the presentation.

--- 2014-04-07 final_presentation/presentation.tex
@@
 \title{Research?}
-\author{Pieter Brederode, Timo Koppenberg, \\ Jelco Bodewes, Anton Golov}
+\author{Pieter Brederode, Jelco Bodewes, \\ Timo Koppenberg, Anton Golov}
 \institute{B3OMI}
@@
     \begin{itemize}
-        \item We did research!
-        \item That's it, really.
+        \item Problem Statement
+        \item Measurement Implementation
+        \item Results
+        \item Conclusions
     \end{itemize}
+\end{frame}
+
+\begin{frame}{Context: Whitelist/Blacklist}
+    \begin{itemize}
+        \item Program that given a URL decides whether to allow access.
+        \item Different possible granualities:
+        \begin{itemize}
+            \item IP address (32 bit integer)
+            \item Domain name (20 char string)
+            \item Full path (100 char string)
+        \end{itemize}
+        \item Fast insert and query
+        \item Low memory usage
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Implementation}
+    \begin{itemize}
+        \item C++
+        \begin{itemize}
+            \item We hoped for predictable performance\ldots
+            \item Standard library has hash table and binary search tree
+            \item Standard library has timing tools
+        \end{itemize}
+        \item Open source SkipList library (\texttt{CSSkipList})
+        \begin{itemize}
+            \item Weird results can be investigated directly
+            \item Turned out useful
+        \end{itemize}
+        \item POSIX API for memory usage
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Implementation Issues}
+    \begin{itemize}
+        \item POSIX API for memory is too coarse
+        \begin{itemize}
+            \item Added tracking of alloctions
+            \item Bonus: only track the data structure itself
+            \item Required modifying the SkipList
+        \end{itemize}
+        \item Performance is consistent, but unusual
+        \begin{itemize}
+            \item Hash tables are quirky
+            \item Suspect: the allocator
+            \item Run tests in separate process
+            \begin{itemize}
+                \item Or write your own allocator\ldots
+            \end{itemize}
+        \end{itemize}
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Small hash tables are expensive}
+
+\end{frame}
+
+\begin{frame}{Big hash tables are cheap}
+
 \end{frame}

### 2014-04-09 final_presentation/presentation.tex

--- 2014-04-09 final_presentation/presentation.tex
@@
     \begin{itemize}
-        \item We did research!
-        \item That's it, really.
+        \item Problem Statement
+        \item Measurement Implementation
+        \item Results
+        \item Conclusions
     \end{itemize}
+\end{frame}
+
+\begin{frame}{Context: Whitelist/Blacklist}
+    \begin{itemize}
+        \item Program that given a URL decides whether to allow access.
+        \item Different possible granualities:
+        \begin{itemize}
+            \item IP address (32 bit integer)
+            \item Domain name (20 char string)
+            \item Full path (100 char string)
+        \end{itemize}
+        \item Fast insert and query
+        \item Low memory usage
+    \end{itemize}
+\end{frame}
+
+\section{Implementation}
+\begin{frame}{Implementation}
+    \begin{itemize}
+        \item C++
+        \begin{itemize}
+            \item We hoped for predictable performance\ldots
+            \item Standard library has hash table and binary search tree
+            \item Standard library has timing tools
+        \end{itemize}
+        \item Open source SkipList library (\texttt{CSSkipList})
+        \begin{itemize}
+            \item Weird results can be investigated directly
+            \item Turned out useful
+        \end{itemize}
+        \item POSIX API for memory usage
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Implementation Issues}
+    \begin{itemize}
+        \item POSIX API for memory is too coarse
+        \begin{itemize}
+            \item Added tracking of alloctions
+            \item Bonus: only track the data structure itself
+            \item Required modifying the SkipList
+        \end{itemize}
+        \item Performance is consistent, but unusual
+        \begin{itemize}
+            \item Hash tables are quirky
+            \item Suspect: the allocator
+            \item Run tests in separate process
+            \begin{itemize}
+                \item Or write your own allocator\ldots
+            \end{itemize}
+        \end{itemize}
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Small hash tables are expensive}
+    \centering
+    \includegraphics[width=0.8\textwidth]{hashtable_bad.png}
+\end{frame}
+
+\begin{frame}{Big hash tables are cheap}
+    \centering
+    \includegraphics[width=0.8\textwidth]{hashtable_good.png}
 \end{frame}

### 2014-04-09 final_presentation/presentation.tex
More presentation work
Fixed some English.

--- 2014-04-09 final_presentation/presentation.tex
@@
 \usenavigationsymbolstemplate{}
-\title{Research?}
+\title{Comparison of Associative Data Structures}
 \author{Pieter Brederode, Jelco Bodewes, \\ Timo Koppenberg, Anton Golov}
@@
 
+\section{Implementation}
 \begin{frame}{Implementation}
@@
 \begin{frame}{Small hash tables are expensive}
-
+    \centering
+    \includegraphics[width=0.8\textwidth]{hashtable_bad.png}
 \end{frame}
@@
 \begin{frame}{Big hash tables are cheap}
+    \centering
+    \includegraphics[width=0.8\textwidth]{hashtable_good.png}
+\end{frame}
 
+\section{Results}
+
+\begin{frame}{Results}
+    \begin{itemize}
+        \item 117 different setups.
+        \item 19 runs, for a total of 2223 cases.
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Number of entries}
+    \begin{figure}
+      \centering
+        \includegraphics[width=0.95\textwidth]{SkiplistQuery}
+      \caption{The average duration of each query operation, plotted against the total number of entries.}
+    \end{figure}
+\end{frame}
+
+\begin{frame}{Number of entries}
+    \begin{itemize}
+        \item There is a positive correlation between the number of entries and the average duration of all operations.
+        \item This is true for both insert and query average duration.
+        \item Not always true for average memory usage.
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Number of entries}
+    \begin{itemize}
+        \item Average memory usage for BST is constant.
+        \item Average memory usage for hash table has a positive correlation with the number of entries.
+        \item Average memory usage for Skiplist has a negative correlation with the number of entries.
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Number of entries}
+    \begin{figure}
+      \centering
+        \includegraphics[width=0.95\textwidth]{SkiplistMemory}
+      \caption{The average duration of each query operation, plotted against the total number of entries.}
+    \end{figure}
+\end{frame}
+
+\begin{frame}{Binary search tree}
+    \begin{itemize}
+        \item For IP adresses: Significantly faster insertions than both other structures for low number of entries.
+        \item Scales badly compared to the other structures.
+        \item Other keys: Always slower than a hash table.
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Hash table}
+    \begin{itemize}
+        \item Significantly higher query speed in all situations compared to other structures.
+        \item Faster insert speed as well in most situations.
+        \item Highest memory usage in all cases.
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Skiplist}
+    \begin{itemize}
+        \item In no tested situations the fasted structure.
+        \item Similar performance to a BST for large numbers of entries.
+        \item Always the lowest memory usage, average memory usage goes down with more entries.
+    \end{itemize}
 \end{frame}
@@
 
-\begin{frame}{Results}
+\begin{frame}{Conclusions}
     \begin{itemize}
-        \item What results?
-        \item What are you talking about?
+        \item Hash table usually the best option.
+        \item Binary search tree better for small numbers of simple keys.
+        \item Skiplist useful for low memory usage.
     \end{itemize}

### 2014-04-09 final_presentation/presentation.tex
Last minor changes.

@@
     \centering
-    \includegraphics[width=0.8\textwidth]{hashtable_bad.png}
+    \includegraphics[width=\textwidth]{hashtable_bad.png}
 \end{frame}
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{hashtable_good.png}
+    \includegraphics[width=\textwidth]{hashtable_good.png}
 \end{frame}
@@
       \centering
-        \includegraphics[width=0.95\textwidth]{SkiplistMemory}
-      \caption{The average duration of each query operation, plotted against the total number of entries.}
+        \includegraphics[width=0.82\textwidth]{SkiplistMemory}
+      \caption{The average memory usage per entry, plotted against the total number of entries.}
     \end{figure}
@@
     \begin{itemize}
-        \item For IP adresses: Significantly faster insertions than both other structures for low number of entries.
+        \item For IP adresses: Significantly faster insertions than both other structures for low numbers of entries.
         \item Scales badly compared to the other structures.

### 2014-04-12 paper/paper.tex
Wrote some very basic stubs.
They may be a tad too informal.

@@
 
-    Ph'nglui mglw'nafh Cthulhu R'lyeh wgah'nagl fhtagn~\cite{CallOfCthulhu}. Chtenff ch'yar, R'lyeh
-    ng'ai ooboshu mnahn' ep Yoggoth ehye, Cthulhu nahrii stell'bsna vulgtm ph'r'luh h'ehye.
-    Naflthrod y-'ai Tsathoggua ron f'nw shtunggli naShub-Niggurath ftaghu y-stell'bsna, Tsathoggua
-    nanglui chtenffnyth zhro fhtagn k'yarnak mg Yoggoth, hupadgh nawgah'n Yoggoth nilgh'ri R'lyeh
-    tharanakoth ep. 'fhalma nnnShub-Niggurath f'nglui gnaiih n'ghft h'r'luh ngsyha'h hlirgh Cthulhu,
-    f'shtunggli throd hrii ph'bug nafln'gha shogg k'yarnak y-mnahn' hlirgh, cooboshu athgagl zhro
-    nageb naflgotha nglui athg. Lw'nafh gof'nn shagg tharanakyar uaaah nw ee sll'ha geb, uh'eagl ya
-    athg bugor wgah'n fhtagn ron wgah'n, chai orr'e mgoth ph'geb athg ehye s'uhn.
+    We did research about what data structure could best be used to represent a blacklist in a 
+    web-filter application.  We compared the insertion speed, query speed, and memory usage of three
+    data structures: the binary search tree, the hash table, and the skip list.  We did this because
+    we really liked these three structures.
+
+	\section{Implementation}
+
+	We used C++ as the implementation language.  libstdc++'s implementations of the standard
+	class templates \texttt{std::map} and \texttt{std::unordered\_map} were used for the binary search
+	tree and the hash table respectively.  As the C++ standard library lacks a skip list\footnote{reference to
+	the standard here}, we used an open-source third party library, \texttt{CSSkipList}\footnote{reference to
+	library home page here}.
+
+	For our measurements, we used libstdc++'s implementation of the standard
+	\texttt{std::chrono::high\_resolution\_clock} class was used for time measurements.   Memory
+	measurements were performed by providing a custom allocator which tracked the number of memory
+	requests.  Note that the latter is a step away from our initial plan, where we stated we would use
+	the POSIX API\footnote{ref to the posix api}; unfortunately, that turned out to be too coarse-grained
+	for our purposes and we could thus get no accurate measurements that way.  The allocator-based technique
+	we employed required slight modifications to the skip list implementation we used, but this
+	shouldn't have caused a significant slowdown (if anything, it made things more fair, as now allocation
+	was done the same way in both systems).
 
@@
 
-    Ilyaa lw'nafh f'sll'ha mg hupadgh li'heeog n'gha athg lloignyth geb, Dagonnyth 'bthnk r'luh
-    sll'ha naflgrah'n gnaiih Shub-Niggurath Tsathoggua stell'bsna f'tharanak, sll'ha nw ep
-    ngstell'bsna f'zhro naflgotha y'hah 'ai. Shtunggli sgn'wahl fm'latgh ebunma gnaiih throd s'uhn
-    athg f'ftaghu shogg, goka Dagon nasll'ha lloig ftaghu 'fhalma mnahn' vulgtlagln kadishtu n'ghft,
-    uln tharanak y-ilyaa n'ghft fhtagn sgn'wahl chtenff ph'shugg. Ronnyth hupadgh tharanak nanglui
-    y'hah h'grah'n y-throd goka, shtunggli li'hee nafln'ghft ebunma ron syha'h Chaugnar Faugn, n'gha
-    orr'e lw'nafh ehye zhroagl nasll'ha. Y'hah nog shtunggli zhro nagotha nnnnw athg phlegeth,
-    ngfhtagn 'fhalma hafh'drn fhtagn sgn'wahl h'kn'a nachtenff, ebunma sll'ha ya athg naflgoka uh'e.
-
-    K'yarnak kadishtu shogg li'heeyar fm'latgh zhro uh'e phlegeth 'bthnk goka fhtagn geb Azathoth
-    geb, Shub-Niggurath ftaghu zhro shogg vulgtlagln ya nnnchtenff gof'nn orr'e nnnshtunggli ron.
-    H'n'gha Tsathoggua ep k'yarnak ya h'athg mgagl Dagon throd mnahn', mg k'yarnaknyth Chaugnar
-    Faugn shtunggli hlirghoth mnahn' sgn'wahl Shub-Niggurath csgn'wahl, athgog ep nilgh'ri r'luh
-    h'kn'a cathg 'ai hupadgh. Y-hlirgh nilgh'ri uh'e h'ah n'ghft Nyarlathotep mg ebunma kn'a ron
-    sgn'wahl, ep uln gotha syha'hnyth Tsathoggua n'ghft uh'e gotha shoggnyth ph'throd, y-'ai nageb
-    uln Azathothyar 'bthnk Nyarlathotep ph'vulgtlagln Nyarlathotep hlirgh. Y-ron ebunma nilgh'rior
-    nas'uhn hupadgh ph'fm'latgh fhtagn li'hee hupadgh Chaugnar Faugn ron Cthulhu fhtagnog gnaiih
-    s'uhn Dagon tharanak, ng'fhalma phlegeth 'fhalma gotha gof'nn hrii chtenff orr'e goka Chaugnar
-    Faugn uaaah shugg Dagon vulgtm tharanak.
-
-    Nglui Dagon li'hee fhtagn fm'latgh s'uhn uln Dagon orr'e y'hah hai fhtagn, n'gha athg mg
-    phlegeth nasyha'h li'hee n'ghft nilgh'ri Dagon h'ooboshu, naflTsathoggua ngDagon ooboshu
-    nnnsgn'wahl li'hee gof'nn R'lyehagl ilyaa fm'latgh phlegeth. Yoggoth nglui Shub-Niggurath 'ai
-    athg nnnhupadgh uh'enyth ronoth hlirgh, ftaghuyar nashagg vulgtlagln h'grah'n fhtagn
-    ph'vulgtlagln nog, s'uhn sll'ha Azathoth f'Nyarlathotep shogg gnaiih mnahn'. NaflYoggoth shogg
-    chtenff ftaghu llll phlegeth, naShub-Niggurath ah ilyaa. Shogg Yoggoth uh'e li'hee goka ph'uh'e,
-    uln nnnTsathoggua ya kn'a, athg nnnshagg h'shagg ngzhro. Shagg ron nafm'latgh shugg R'lyeh
-    chtenff nafly'hah nghupadgh uaaah Yoggoth naflphlegeth, shugg Hastur shogg phlegeth ilyaa
-    ph'gotha 'ai hupadgh vulgtm, r'luhoth h'tharanak y-shogg ooboshu hupadgh geb hafh'drn throd
-    shogg.
+	Our results were great!  They are best analysed with a two-sample independent T test.  We had an
+	equal number of measurements about group 1 (with mean $x_1$ and standard deviation $\sigma_1$ and
+	about group 2 (with mean $x_2$ and standard deviation $\sigma_2$).  From this we can conclude that
+	data structure 1 is, on average, unquestionably superior.
 
@@
 
-    Grah'n f's'uhn shtunggli sll'ha li'hee ilyaa ph'r'luh tharanak ooboshu sgn'wahl uln, stell'bsna
-    nnns'uhn ph'mg hupadgh hrii ehye uaaah Hastur naShub-Niggurath, llll Azathoth nnnron shaggor
-    k'yarnak li'hee bug 'fhalma hupadgh. Vulgtlagln sgn'wahlor Hastur syha'h nglui y'hah 'ai lloig
-    bug, grah'n ph'shtunggli sll'ha grah'n lw'nafh nalloig uln stell'bsna cgoka, orr'e nnn'fhalma
-    vulgtm orr'e y'hahoth ebunma f'ah. Ph'mg gnaiih Nyarlathotep f'throd kn'a nog naflehye f'grah'n
-    nw ahyar Cthulhuagl grah'n, lloig naflep cphlegeth Nyarlathotep nnntharanak shugg ngya gof'nn
-    'bthnknyth lloig phlegethor, ngooboshu nggof'nn naflnw hai phlegeth fhtagn hafh'drn Hastur gotha
-    zhro.
+    The skip list uses the least memory but is awful in every other imaginable way.  The hashmap is
+    usually the best, except when we're dealing with a small number of integer keys and doing far more
+    inserts than queries.  Note that this doesn't actually happen in practice.
 

### 2014-04-14 paper/paper.tex
Made a bunch of notes about implementation issues.

@@
     \begin{abstract}
-        We did research and got results.
+        We did research and got results~\cite{CallOfCthulhu}.
     \end{abstract}
@@
 
+    \subsection{Possible errors caused by the implementation being not entirely perfect}
+
+    We really did our best to make sure these kinds of things didn't happen.  However, there were
+    still a number of measurements that do not follow from the underlying theory and are likely a
+    consequence of the hardware involved.
+
+    In particular, constructing a small hash table after constructing many big ones takes an
+    absurdly long time.  Only the first hash table is affected, and the time is comparable to the
+    time required to construct a 4k element hash table, which is just silly.  It probably has
+    something to do with the allocator but the reason for such behaviour is still unclear.  The fact
+    a hashtable needs contiguous memory probably has something to do with it, as well as the
+    allocator returning memory to the operating system.  Still, the latter should also have happened
+    with the other structures.  Bizarre.
+
+    A similar issue occurs around 4k or 8k (depending on data structure).  In all three cases, the
+    cost of insertion and querying grows at a higher rate than previously.  This is difficult to
+    quantify\footnote{Can we do some sort of linear regression to prove this is the case?}, but it
+    seems like some extra costs are added around then.  We suspect it is related either to the
+    allocator having to get more memory or the CPU cache running out and more cache misses occuring.
+    We'd measure that too if we had more time/resources.
+
+    A possibly related issue is that the standard deviation of our measurements rises around that
+    point.  That strengthens the possibility of it being allocator-related, as that would involve a
+    system call and could cause scheduling issues (despite the system being otherwise very close to
+    idle).
+
     \section{Conclusions}

### 2014-04-15 paper/paper.tex
Wrote some more on the paper.

@@
 
-    We used C++ as the implementation language.  libstdc++'s implementations of the standard
-    class templates \texttt{std::map} and \texttt{std::unordered\_map} were used for the binary search
-    tree and the hash table respectively.  As the C++ standard library lacks a skip list\footnote{reference to
-    the standard here}, we used an open-source third party library, \texttt{CSSkipList}\footnote{reference to
-    library home page here}.
+    We used C++ as the implementation language.  libstdc++'s implementations of the standard class
+    templates \texttt{std::map} and \texttt{std::unordered\_map} were used for the binary search
+    tree and the hash table respectively.  As the C++ standard library lacks a skip
+    list\footnote{reference to the standard here}, we used an open-source third party library,
+    \texttt{CSSkipList}\footnote{reference to library home page here}.  We also used libstdc++'s
+    implementation of \texttt{std::string} for our string handling needs.
 
@@
     measurements were performed by providing a custom allocator which tracked the number of memory
-    requests.  Note that the latter is a step away from our initial plan, where we stated we would use
-    the POSIX API\footnote{ref to the posix api}; unfortunately, that turned out to be too coarse-grained
-    for our purposes and we could thus get no accurate measurements that way.  The allocator-based technique
-    we employed required slight modifications to the skip list implementation we used, but this
-    shouldn't have caused a significant slowdown (if anything, it made things more fair, as now allocation
-    was done the same way in both systems).
+    requests.  Note that the latter is a step away from our initial plan, where we stated we would
+    use the POSIX API\footnote{ref to the posix api}; unfortunately, that turned out to be too
+    coarse-grained for our purposes and we could thus get no accurate measurements that way.  The
+    allocator-based technique we employed required slight modifications to the skip list
+    implementation we used, but this shouldn't have caused a significant slowdown (if anything, it
+    made things more fair, as now allocation was done the same way in both systems).
+
+    For our data, we took the system's dict file and the list of top level domains from Wikipedia
+    and combined the two to generate a bunch of things of the form \texttt{www.\$word.\$tld}, where
+    \texttt{\$word} is one of the dictionary words and \texttt{\$tld} is a top level domain.  Only
+    words consisting of the lowercase letters \texttt{a} through \texttt{z} were used.  Words were
+    limited to be from 8 to 12 characters and top-level domains were limited to exactly two letters.
+    All possible combinations were generated.  For the full paths, a number of domains were selected
+    and each was appended some number of random alphanumeric strings of eighty characters.  For more
+    details, see the code.
+
+    After some test runs we chose to decrease the number of runs used for measurements.  In the
+    proposal we had said that for each time measurement we would perform the relevant operation 1000
+    times and then take the average of that.  In practice, this made the tests run for longer than
+    we could afford and the results given by running 100 times were comparable so we only ran that
+    many times.  We expect the general result to be the same; see some later section for details
+    about what \emph{could} have gone wrong.
 
@@
 
+    Also, we have counterintuitive results that show that inserting or querying a full path is
+    faster than querying only the domain name.  This can partially be explained\footnote{Not really,
+    but ehhh.} by the fact that the string implementation we used was based on reference counting.
+    (If it'd also provide SSO, this would be easy to explain, but the darn thing doesn't.)
+
     \section{Conclusions}

### 2014-04-15 paper/paper.tex
rewrote the implementation section

--- 2014-04-14 paper/paper.tex
@@
     \section{Implementation}
+    \subsubsection*{Datastructures}
 
-    We used C++ as the implementation language.  libstdc++'s implementations of the standard
-    class templates \texttt{std::map} and \texttt{std::unordered\_map} were used for the binary search
-    tree and the hash table respectively.  As the C++ standard library lacks a skip list\footnote{reference to
-    the standard here}, we used an open-source third party library, \texttt{CSSkipList}\footnote{reference to
-    library home page here}.
+    Using C++ as the implementation language, we chose libstdc++'s implementations of the standard class
+    templates \texttt{std::map} and \texttt{std::unordered\_map} for the binary search
+    tree and the hash table respectively.  As the C++ standard library does not contain a skip
+    list\footnote{reference to the standard here}, we used an open-source third party library,
+    \texttt{CSSkipList}\footnote{reference to library home page here}.  We also used libstdc++'s
+    implementation of \texttt{std::string} for our string handling\footnote{specify?}needs.
 
-    For our measurements, we used libstdc++'s implementation of the standard
-    \texttt{std::chrono::high\_resolution\_clock} class was used for time measurements.   Memory
-    measurements were performed by providing a custom allocator which tracked the number of memory
-    requests.  Note that the latter is a step away from our initial plan, where we stated we would use
-    the POSIX API\footnote{ref to the posix api}; unfortunately, that turned out to be too coarse-grained
-    for our purposes and we could thus get no accurate measurements that way.  The allocator-based technique
-    we employed required slight modifications to the skip list implementation we used, but this
-    shouldn't have caused a significant slowdown (if anything, it made things more fair, as now allocation
-    was done the same way in both systems).
+    \subsubsection*{Data}
+    For our data, we took the system's dictionary file and the list of top level domains from Wikipedia
+    and combined these to generate a list of items in the form \texttt{www.\$word.\$tld}, where
+    \texttt{\$word} is one of words from the the dictionary and \texttt{\$tld} is a top level domain.  Only
+    words consisting of the lowercase letters \texttt{a} through \texttt{z} were used.  Constraining the
+    words to consist of 8 to 12 characters and limiting the top-level domains to exactly two letters, 
+    all possible combinations were generated.  For the full paths, a number of domains was selected
+    and each was appended a random alphanumeric string of eighty characters.  The exact implementation
+    can be viewed in the sourcecode\footnote{ref to apendix or code dump}.
+
+	\subsubsection*{Measurements}
+    For our time measurements, we used libstdc++'s implementation of the standard
+    \texttt{std::chrono::high\_resolution\_clock} class.
+
+    Memory measurements were performed by providing a custom allocator which tracked the number of memory
+    requests.  Note that the latter deviates from our initial plan, where we stated\footnote{ref?}we would
+    use the POSIX API\footnote{ref to the posix api}.  Unfortunately, that turned out to be too
+    coarse-grained for our purposes and we could thus get no accurate measurements.  The
+    allocator-based technique we employed did require some slight modifications to the skip list
+    implementation we used, but this should not have caused a significant slowdown. We assume that any 
+    influence will only have made the tests more fair, as now allocation was done the same way in all 
+    three setups.
+
+    After some tests we chose to decrease the number of runs used for measurements.  In the
+    proposal we stated that for each time measurement we would perform the relevant operation 1000
+    times and then take the average of that.  In practice, this made the tests run for longer than
+    we could afford. Because the results given by performing 100 runs were comparable, we settled for that.  
+    We expect the general result to be similar; refer to section\footnote{ref}for details
+    about what \emph{could} have gone wrong.
 

### 2014-04-16 paper/paper.tex
minor change
my dad teaches english and corrected me here (on my request)
Wrote the introduction and started on the conclusion.

@@
     \section{Introduction}
+    A website filter is a program that catagorises websites in order to restrict access to designated
+    sites a user might visit. Commonly, the filter includes a blacklist of sites a user may not visit
+    unless they have certain credentials, or alternatively, includes a whitelist so it can blacklist
+    everything else. As the world-wide web is growing, these lists have grown larger to a point where
+    the speed of the underlying data structure has become a key issue. The time of both inserting and
+    looking up websites must be kept minimal. This all must be done without too much of a trade-off
+    between speed and memory usage.
 
-    We did research about what data structure could best be used to represent a blacklist in a 
-    web-filter application.  We compared the insertion speed, query speed, and memory usage of three
-    data structures: the binary search tree, the hash table, and the skip list.  We did this because
-    we really liked these three structures.
+    In technical terms, we would want the best data structure for associating many URLs with their
+    respective permissions, be it whitelisted or blacklisted for certain users. In researching the best
+    associative data structure we have picked implementations of the binary search tree, the hash map
+    and the skip list. These are the key options in such a problem. We are interested in finding which
+    one would perform best under a various number of insertions. The main research question of this paper
+    is as follows:
 
+    How does the number of entries influence the performance of various associative data structures?
+
+    While the decision of data structure might mostly be dependant on the number of entries, the complexity
+    of the key might also be of concern. Where permissions per page would be preferred, a more
+    resource optimal solution might be found in filtering only on domain or even IP adress.
+    In researching the influence of this we formulate a secondary research question:
+
+    How does the complexity of the key influence the preformance of various associative data structures?
+
+    In the following paper, we will attempt to answer these questions. The next section will give details
+    about our research plan and the scopes of our research. The section thereafter is about the implementation
+    of our setup in C++. Next, we will state our results, after which we will discuss set results and provide
+    several conclusions.    
+    %We did research about what data structure could best be used to represent a blacklist in a 
+    %web-filter application.  We compared the insertion speed, query speed, and memory usage of three
+    %data structures: the binary search tree, the hash table, and the skip list.  We did this because
+    %we really liked these three structures. 
     \section{Implementation}
@@
     words to consist of 8 to 12 characters and limiting the top-level domains to exactly two letters, 
-    all possible combinations were generated.  For the full paths, a number of domains was selected
+    all possible combinations were generated.  For the full paths, a number of domains were selected
     and each was appended a random alphanumeric string of eighty characters.  The exact implementation
@@
     \section{Conclusions}
+    Our main point of research was researching which associative data structure performed best on
+    a various number of insertions, in terms of ths speed of insertions and look-ups, and memory.
+    In terms of speed we can state with certainty that the hash table is the superior data structure.
+    It outperforms the other two in almost every category, most significantly in larger datastructures where
+    it is often takes less than half as much time as the skiplist and less than twothirds the time of the binary search tree.
+    In most cases, the binary search tree is a bit slower, where the skip list is the slowest structure in
+    virtually every case. The only motivation for using a hashmap for this problem appears to be in memory usage,
+    although the advantage gained from this is quite small.
 

### 2014-04-17 paper/paper.tex
Wrote the conclusion and reflection

@@
 
-    The skip list uses the least memory but is awful in every other imaginable way.  The hashmap is
-    usually the best, except when we're dealing with a small number of integer keys and doing far more
-    inserts than queries.  Note that this doesn't actually happen in practice.
+    Our secondary point of research encompassed the effect of key complexity on associative data structures.
+    Our different types of keys were IP adresses, domain names and full path URLs. We have seen that as expected,
+    IP adresses as keys give a significant speedup over the other two, and use less memory as well. Interesting to
+    note is that we found the only advantage of the binary search tree over the hashmap here, in that insertions of
+    a small amount of IP adresses, up to around $2^13$ insertions, is faster. So when writing an application with very
+    few keys of low complexity, with more insertions than queries, the binary search tree may be the best option. Lastly,
+    we have found that the domain key type does not outperform the full path key type. It is not significantly better
+    if it is better at all.
+    %The skip list uses the least memory but is awful in every other imaginable way.  The hashmap is
+    %usually the best, except when we're dealing with a small number of integer keys and doing far more
+    %inserts than queries.  Note that this doesn't actually happen in practice.
+
+    \section{Reflection}
+    In general, we are glad with the results of our research. We were able to provide clear and concise answers to our
+    research questions. Even though our research questions may appear broad, we have defined a good scope in our research
+    proposal and have virtually not mended this scope in our research. One small change is that we also included datasets of
+    size $2^8$ because we saw some interesting results in very small sets. In our implementation we have changed quite a few
+    things because we ran into pratical problems. Some of these required us to change the allocator and generate our keys
+    in a slightly different fashion, which we all documented in this paper.
+
+    Since we have found that the hash map is the best associative data structure overal for this problem, we might do further
+    research in different kind of hash maps. We might try different hash functions, different methods of collision resolution,
+    and perhaps survey different kinds of concurrent hash maps.
 

### 2014-04-17 paper/paper.tex
wrote the abstract

@@
 
-\title{Data Structures Comparative Review Thing}
+\title{Associative Data Structures for Website Filtering}
 \author{Pieter Brederode, Timo Koppenberg, \\ Jelco Bodewes, Anton Golov}
@@
     \begin{abstract}
-        We did research and got results~\cite{CallOfCthulhu}.
+        In this paper, we will research associative data structures. This is done in the context of
+        a website filter. We look at three different data structures, the hash map, binary search tree
+        and the skip list. We insert keys of three different types, namely IP adress, domain name and
+        full path URL. We judge the competency of our data strucutre based on insert time, query time
+        and memory usage, inserting data sets of sizes $2^x$ for $8 \le x \le 20$. We describe the details
+        of our research and the implementation using C++. The results are that hash maps are much faster
+        than the other structures, where skip lists use the least memory and binary search trees are
+        only useful with very few insertions of simple keys and more insertions than queries.
     \end{abstract}

### 2014-04-17 paper/paper.tex
wrote the abstract, added quotues to the research questions

@@
 
-    How does the number of entries influence the performance of various associative data structures?
+    \begin{quotation}How does the number of entries influence the performance of various associative data structures?\end{quotation}
 
@@
 
-    How does the complexity of the key influence the preformance of various associative data structures?
+    \begin{quotation}How does the complexity of the key influence the preformance of various associative data structures?\end{quotation}
 

### 2014-04-17 paper/paper.tex
implementation errors, typos

@@
 \title{Associative Data Structures for Website Filtering}
-\author{Pieter Brederode, Timo Koppenberg, \\ Jelco Bodewes, Anton Golov}
+\author{Jelco Bodewes, Pieter Brederode \\ Anton Golov, Timo Koppenberg }
 
@@
     \begin{abstract}
-        In this paper, we will research associative data structures. This is done in the context of
-        a website filter. We look at three different data structures, the hash map, binary search tree
+        In this paper, we will research associative data structures for searching. This is done in the context of
+        a website filter. We look at three different data structures: the hash map, the binary search tree
         and the skip list. We insert keys of three different types, namely IP adress, domain name and
-        full path URL. We judge the competency of our data strucutre based on insert time, query time
+        full path URL. We judge the competency of our data strucutre based on insertion times, query times
         and memory usage, inserting data sets of sizes $2^x$ for $8 \le x \le 20$. We describe the details
         of our research and the implementation using C++. The results are that hash maps are much faster
-        than the other structures, where skip lists use the least memory and binary search trees are
+        than the other structures, skip lists use the least memory and binary search trees are
         only useful with very few insertions of simple keys and more insertions than queries.
@@
     \section{Introduction}
-    A website filter is a program that catagorises websites in order to restrict access to designated
+    A website filter is a program that categorises websites in order to restrict access to designated
     sites a user might visit. Commonly, the filter includes a blacklist of sites a user may not visit
@@
 
-    In the following paper, we will attempt to answer these questions. The next section will give details
+    In the following paper, we will attempt to answer these questions. The next section will provide details
     about our research plan and the scopes of our research. The section thereafter is about the implementation
@@
 
-    \subsection{Possible errors caused by the implementation being not entirely perfect}
+    \subsection{Errors possibly caused by an imperfect implementation}
 
-    We really did our best to make sure these kinds of things didn't happen.  However, there were
-    still a number of measurements that do not follow from the underlying theory and are likely a
+    There were a number of measurements that did not follow the underlying theory and are likely a
     consequence of the hardware involved.
 
-    In particular, constructing a small hash table after constructing many big ones takes an
-    absurdly long time.  Only the first hash table is affected, and the time is comparable to the
-    time required to construct a 4k element hash table, which is just silly.  It probably has
-    something to do with the allocator but the reason for such behaviour is still unclear.  The fact
-    a hashtable needs contiguous memory probably has something to do with it, as well as the
-    allocator returning memory to the operating system.  Still, the latter should also have happened
-    with the other structures.  Bizarre.
+    In particular, constructing a small hash table\footnote{state the number of elements in a "small" hash table}
+    after constructing many big ones takes a disproportionally long time.
+Only the first hash table is affected, and the time is 			comparable to the time required to construct a hash table with 4000 elements, which is illogical.
+    We assume this is caused by the allocator but the reason for such behaviour is still unclear.  The fact
+    that a hashtable needs contiguous memory is likely to have some influence, as well as the
+    allocator returning memory to the operating system.  However, we reason the latter should also have happened
+    with the other structures.
 
-    A similar issue occurs around 4k or 8k (depending on data structure).  In all three cases, the
-    cost of insertion and querying grows at a higher rate than previously.  This is difficult to
+    A similar issue occurs around 4000 or 8000 elements (depending on data structure).  For all three
+    datastructures, the cost of inserting and querying grows at a higher rate than before.  This is difficult to
     quantify\footnote{Can we do some sort of linear regression to prove this is the case?}, but it
-    seems like some extra costs are added around then.  We suspect it is related either to the
-    allocator having to get more memory or the CPU cache running out and more cache misses occuring.
-    We'd measure that too if we had more time/resources.
+    appears some extra load is added at that point.  We suspect it is related either to the
+    allocator having to acquire more memory or the CPU cache running out and more cache misses occuring.
+    We would measure this as well, if there was more time.
 
-    A possibly related issue is that the standard deviation of our measurements rises around that
-    point.  That strengthens the possibility of it being allocator-related, as that would involve a
-    system call and could cause scheduling issues (despite the system being otherwise very close to
-    idle).
+    A potentially related issue is that the standard deviation of our measurements rises around the
+    same point.  That increases the likelihood of it being allocator-related, as that involves a
+    system call and could cause scheduling issues\footnote{elaborate/ref?}, despite the system being 
+    otherwise very close to idle.
 
-    Also, we have counterintuitive results that show that inserting or querying a full path is
+    Also, we acquired counterintuitive results showing that inserting or querying a full path is
     faster than querying only the domain name.  This can partially be explained\footnote{Not really,
-    but ehhh.} by the fact that the string implementation we used was based on reference counting.
-    (If it'd also provide SSO, this would be easy to explain, but the darn thing doesn't.)
+    but ehhh.}by the fact that the string implementation we used was based on reference counting.
+    %(If it'd also provide SSO, this would be easy to explain, but the darn thing doesn't.)
 
@@
     Our main point of research was researching which associative data structure performed best on
-    a various number of insertions, in terms of ths speed of insertions and look-ups, and memory.
+    a various number of insertions, in terms of the speed of insertions and look-ups, and memory usage.
     In terms of speed we can state with certainty that the hash table is the superior data structure.
     It outperforms the other two in almost every category, most significantly in larger datastructures where
-    it is often takes less than half as much time as the skiplist and less than twothirds the time of the binary search tree.
-    In most cases, the binary search tree is a bit slower, where the skip list is the slowest structure in
+    it often takes less than half as much time as the skiplist and less than two thirds the time of the binary search tree.
+    In most cases, the binary search tree is a bit slower, whereas the skip list is the slowest structure in
     virtually every case. The only motivation for using a hashmap for this problem appears to be in memory usage,
-    although the advantage gained from this is quite small.
+    although the advantage gained from this is relatively small.
 
@@
     Our different types of keys were IP adresses, domain names and full path URLs. We have seen that as expected,
-    IP adresses as keys give a significant speedup over the other two, and use less memory as well. Interesting to
-    note is that we found the only advantage of the binary search tree over the hashmap here, in that insertions of
+    IP adresses as keys provide a significant improvement over the other two, and uses less memory as well. Interesting to
+    note is that we found the only advantage of the binary search tree over the hashmap here, is that insertions of
     a small amount of IP adresses, up to around $2^13$ insertions, is faster. So when writing an application with very
-    few keys of low complexity, with more insertions than queries, the binary search tree may be the best option. Lastly,
+    few keys of low complexity, and with more insertions than queries, the binary search tree may be the best option. Finally,
     we have found that the domain key type does not outperform the full path key type. It is not significantly better
@@
     \section{Reflection}
-    In general, we are glad with the results of our research. We were able to provide clear and concise answers to our
+    In general, we are content with the results of our research. We were able to provide clear and concise answers to our
     research questions. Even though our research questions may appear broad, we have defined a good scope in our research
-    proposal and have virtually not mended this scope in our research. One small change is that we also included datasets of
-    size $2^8$ because we saw some interesting results in very small sets. In our implementation we have changed quite a few
-    things because we ran into pratical problems. Some of these required us to change the allocator and generate our keys
+    proposal and have virtually not altered this scope in our research. One small change is that we also included datasets of
+    size $2^8$ because we saw some interesting results in very small sets. In our implementation we have made some significant
+    changes because we ran into pratical problems. Some of these required us to change the allocator and generate our keys
     in a slightly different fashion, which we all documented in this paper.
 
-    Since we have found that the hash map is the best associative data structure overal for this problem, we might do further
+    Since we have found that the hash map is the best associative data structure overal for this problem, we could do further
     research in different kind of hash maps. We might try different hash functions, different methods of collision resolution,

### 2014-04-17 paper/paper.tex
research description
ik heb dit gemerged met het implementation deel wat er al stond, er zat
erg veel overlap in

@@
 
-    \begin{quotation}How does the number of entries influence the performance of various associative data structures?\end{quotation}
+    \emph{How does the number of entries influence the performance of various associative data structures?}
 
@@
 
-    \begin{quotation}How does the complexity of the key influence the preformance of various associative data structures?\end{quotation}
+    \emph{How does the complexity of the key influence the preformance of various associative data structures?}
 
-    In the following paper, we will attempt to answer these questions. The next section will provide details
-    about our research plan and the scopes of our research. The section thereafter is about the implementation
-    of our setup in C++. Next, we will state our results, after which we will discuss set results and provide
-    several conclusions.    
+    In this paper, we will attempt to answer these questions. The next section will provide details
+    about our research plan (based largely on our original plan) and the implementation of our setup in C++. 	 
+    Next, we will state our results, after which we will discuss set results and provide several conclusions.    
     %We did research about what data structure could best be used to represent a blacklist in a 
@@
     %we really liked these three structures.
-    \section{Implementation}
+
+    \section{Research description}
+
+    \subsubsection*{Scope and assumptions}
+
+    We make several assumptions for our research.
+
+	Firstly we plan on only working with datasets that are certain to fit into the main memory of our machine. We intend to 
+    purely examine the relative speed and memory-efficiency of the different data structures. Using datasets larger than 
+    memory will cause regular lookups to the harddrive, which we assume would skew our results.
+
+    In order to make the measurements more predictable, we will always ensure that insertions are only performed with 
+    non-existent keys. The alternative would make either the memory usage or number of insertions non-deterministic, which 
+    could cause inaccuracies in measurement. Similarly, we require that all queries be performed on existing keys.
+
+	To mimic real usage, we will insert and then access the keys in an arbitrary order.  However, we will keep the order consistent 
+    between tests on different data structures.  Together this will minimize any bias related to any kind of order-dependent 
+    performance, which would likely be an advantage specific to the implementation, not to the data structure. We will 
+    generate the order for both lookup and insertion by shuffling the dataset we chose for the specific test. This means that 
+    orders will not be persistent between different data set sizes; however, we do not expect this to be significant.
+
+	We realize that methods to measure the amount of time required for actions will themselves need time. We will however 
+    use the same time-measuring method for all data structures, so we can still use the results to compare the different 
+    data structures.
+
     \subsubsection*{Datastructures}
@@
     \subsubsection*{Data}
-    For our data, we took the system's dictionary file and the list of top level domains from Wikipedia
-    and combined these to generate a list of items in the form \texttt{www.\$word.\$tld}, where
-    \texttt{\$word} is one of words from the the dictionary and \texttt{\$tld} is a top level domain.  Only
-    words consisting of the lowercase letters \texttt{a} through \texttt{z} were used.  Constraining the
-    words to consist of 8 to 12 characters and limiting the top-level domains to exactly two letters, 
-    all possible combinations were generated.  For the full paths, a number of domains were selected
-    and each was appended a random alphanumeric string of eighty characters.  The exact implementation
-    can be viewed in the sourcecode\footnote{ref to apendix or code dump}.
+    For our data, we used datasets containing $2^x$ elements with $x$ an integer such that $8 < x < 21$. Thus we have 
+    twelve different dataset sizes. we took the system's dictionary file and the list of top level domains from Wikipedia 
+    and combined these to generate a list of items in the form \texttt{www.\$word.\$tld}, where \texttt{\$word} is one 
+    of words from the the dictionary and \texttt{\$tld} is a top level domain.  Only words consisting of the lowercase 
+    letters \texttt{a} through \texttt{z} were used. Constraining the words to consist of 8 to 12 characters and limiting 
+    the top-level domains to exactly two letters, all possible combinations were generated.  For the full paths, a number 
+    of domains were selected and each was appended a random alphanumeric string of eighty characters.  Finally, for the 
+    third type of key, an IP address, we used 32 bit integers. The exact implementation can be viewed in the 
+    sourcecode\footnote{ref to apendix or code dump}.
 
@@
     In particular, constructing a small hash table\footnote{state the number of elements in a "small" hash table}
-    after constructing many big ones takes a disproportionally long time.
-Only the first hash table is affected, and the time is 			comparable to the time required to construct a hash table with 4000 elements, which is illogical.
+    after constructing many big ones takes a disproportionally long time.  Only the first hash table is affected, 
+    and the time is comparable to the time required to construct a hash table with 4000 elements, which is illogical.
     We assume this is caused by the allocator but the reason for such behaviour is still unclear.  The fact

### 2014-04-17 paper/paper.tex
Tried to make things more formal.

--- 2014-04-17 paper/paper.tex
@@
     \begin{abstract}
-        In this paper, we will research associative data structures for searching. This is done in the context of
-        a website filter. We look at three different data structures: the hash map, the binary search tree
-        and the skip list. We insert keys of three different types, namely IP adress, domain name and
-        full path URL. We judge the competency of our data strucutre based on insertion times, query times
-        and memory usage, inserting data sets of sizes $2^x$ for $8 \le x \le 20$. We describe the details
-        of our research and the implementation using C++. The results are that hash maps are much faster
-        than the other structures, skip lists use the least memory and binary search trees are
-        only useful with very few insertions of simple keys and more insertions than queries.
+        We have researched the performance of the hash table, binary search tree, and skip list
+        associative datastructures in the context of a web filter.  Our measurements were made with
+        structures of various sizes (ranging from $2^8$ up to $2^{20}$) and with different possible
+        key types (from 4 to 120 bytes).  In this paper, we describe our implementation and
+        empirically show that given a sufficiently high number of entries, a hash table has the
+        highest insertion and query speeds, while the skip list requires the least memory per
+        element.
+        % And when I say "hash table", I mean libstdc++'s unordered_map, and when I say bst I mean
+        % libstdc++'s map, and if you took different implementations you'd probably get different
+        % results...
     \end{abstract}
@@
     A website filter is a program that categorises websites in order to restrict access to designated
-    sites a user might visit. Commonly, the filter includes a blacklist of sites a user may not visit
-    unless they have certain credentials, or alternatively, includes a whitelist so it can blacklist
-    everything else. As the world-wide web is growing, these lists have grown larger to a point where
+    sites a user might visit. Commonly, the filter operates by keeping a blacklist of sites a user may not visit
+    unless they have certain credentials, or alternatively, keeping a whitelist and prohibiting any
+    sites not on it.  As the world-wide web is growing, these lists have grown larger to a point where
     the speed of the underlying data structure has become a key issue.  The time of both inserting and
-    looking up websites must be kept minimal. This all must be done without too much of a trade-off
-    between speed and memory usage.
+    looking up websites must be kept minimal. Moreover, the increased number of entries make memory
+    usage an equally significant concern.
 
-    In technical terms, we would want the best data structure for associating many URLs with their
-    respective permissions, be it whitelisted or blacklisted for certain users. In researching the best
-    associative data structure we have picked implementations of the binary search tree, the hash map
-    and the skip list. These are the key options in such a problem. We are interested in finding which
-    one would perform best under a various number of insertions. The main research question of this paper
-    is as follows:
+    In technical terms, we are interested in the best data structure for associating many URLs with
+    their respective permissions, be it whitelisted or blacklisted for certain users.  We have
+    chosen to compare the performance of the binary search tree, the hash map, and the skip list, as
+    these are commonly employed in practice.  It is of particular interest how each data structure
+    will scale as the number of elements rises.  We have thus chosen for our main research question
+    the following:
+    % Guys, did you notice we did absolutely nothing about predicting this?
 
-    \begin{quotation}How does the number of entries influence the performance of various associative data structures?\end{quotation}
+    \begin{quotation}
+        How does the number of entries influence the performance of various associative data
+        structures?
+    \end{quotation}
 
-    While the decision of data structure might mostly be dependant on the number of entries, the complexity
-    of the key might also be of concern. Where permissions per page would be preferred, a more
-    resource optimal solution might be found in filtering only on domain or even IP adress.
-    In researching the influence of this we formulate a secondary research question:
+    While the decision of data structure might mostly be dependant on the number of entries, the
+    complexity of the key might also be of concern. Where permissions per page would be preferred, a
+    more resource optimal solution might be found in filtering only on domain or even IP adress.  In
+    researching the influence of this we formulate a secondary research question:
 
-    \begin{quotation}How does the complexity of the key influence the preformance of various associative data structures?\end{quotation}
+    \begin{quotation}
+        How does the complexity of the key influence the preformance of various associative data structures?
+    \end{quotation}
 
-    In the following paper, we will attempt to answer these questions. The next section will provide details
-    about our research plan and the scopes of our research. The section thereafter is about the implementation
-    of our setup in C++. Next, we will state our results, after which we will discuss set results and provide
-    several conclusions.    
-    %We did research about what data structure could best be used to represent a blacklist in a 
-    %web-filter application.  We compared the insertion speed, query speed, and memory usage of three
-    %data structures: the binary search tree, the hash table, and the skip list.  We did this because
-    %we really liked these three structures. 
+    In this paper, we attempt to answer these questions. Section~\ref{sec:plan} provides details
+    about our research plan and the scopes of our research. Section~\ref{sec:implementation} is
+    about the implementation of our setup in C++.  The rest of the paper states our results and the
+    conclusions that can be drawn from them.
+
+    \section{Plan}
+    \label{sec:plan}
+
     \section{Implementation}
-    \subsubsection*{Datastructures}
+    \label{sec:implementation}
+
+    We used C++ implementations of the data structures involved, and also the language for our
+    measuring code.  We chose C++ as it is an industry standard\footnote{ref}, provides the
+    necessary tools, and does not carry the risk of garbage collection pauses or effects due to
+    ``warming-up time''.
+
+    \subsection*{Data Structures}
 
@@
 
-    \subsubsection*{Data}
+    \subsection*{Data}
     For our data, we took the system's dictionary file and the list of top level domains from Wikipedia
@@
 
-	\subsubsection*{Measurements}
+    \subsection*{Measurements}
     For our time measurements, we used libstdc++'s implementation of the standard

### 2014-04-17 paper/paper.tex

@@
     \section{Introduction}
-    A website filter is a program that categorises websites in order to restrict access to designated
-    sites a user might visit. Commonly, the filter operates by keeping a blacklist of sites a user may not visit
-    unless they have certain credentials, or alternatively, keeping a whitelist and prohibiting any
-    sites not on it.  As the world-wide web is growing, these lists have grown larger to a point where
-    the speed of the underlying data structure has become a key issue.  The time of both inserting and
-    looking up websites must be kept minimal. Moreover, the increased number of entries make memory
-    usage an equally significant concern.
+    A website filter is a program that categorises websites in order to restrict access to
+    designated sites a user might visit. Commonly, the filter operates by keeping a blacklist of
+    sites a user may not visit unless they have certain credentials, or alternatively, keeping a
+    whitelist and prohibiting any sites not on it.  As the world-wide web is growing, these lists
+    have grown larger to a point where the speed of the underlying data structure has become a key
+    issue.  The time of both inserting and looking up websites must be kept minimal. Moreover, the
+    increased number of entries make memory usage an equally significant concern.
 
@@
 
-    \section{Plan}
+    \section{Research Plan}
     \label{sec:plan}
+
+    \subsubsection*{Scope and assumptions}
+
+    We make several assumptions for our research.
+
+    Firstly we plan on only working with datasets that are certain to fit into the main memory of our machine. We intend to
+    purely examine the relative speed and memory-efficiency of the different data structures. Using datasets larger than
+    memory will cause regular lookups to the harddrive, which we assume would skew our results.
+
+    In order to make the measurements more predictable, we will always ensure that insertions are only performed with
+    non-existent keys. The alternative would make either the memory usage or number of insertions non-deterministic, which
+    could cause inaccuracies in measurement. Similarly, we require that all queries be performed on existing keys.
+
+    To mimic real usage, we will insert and then access the keys in an arbitrary order.  However, we will keep the order consistent
+    between tests on different data structures.  Together this will minimize any bias related to any kind of order-dependent
+    performance, which would likely be an advantage specific to the implementation, not to the data structure. We will
+    generate the order for both lookup and insertion by shuffling the dataset we chose for the specific test. This means that
+    orders will not be persistent between different data set sizes; however, we do not expect this to be significant.
+
+    We realize that methods to measure the amount of time required for actions will themselves need time. We will however
+    use the same time-measuring method for all data structures, so we can still use the results to compare the different
+    data structures.
 
@@
 
-    \subsection*{Data}
-    For our data, we took the system's dictionary file and the list of top level domains from Wikipedia
-    and combined these to generate a list of items in the form \texttt{www.\$word.\$tld}, where
-    \texttt{\$word} is one of words from the the dictionary and \texttt{\$tld} is a top level domain.  Only
-    words consisting of the lowercase letters \texttt{a} through \texttt{z} were used.  Constraining the
-    words to consist of 8 to 12 characters and limiting the top-level domains to exactly two letters, 
-    all possible combinations were generated.  For the full paths, a number of domains were selected
-    and each was appended a random alphanumeric string of eighty characters.  The exact implementation
-    can be viewed in the sourcecode\footnote{ref to apendix or code dump}.
+    \subsubsection*{Data}
+    For our data, we used datasets containing $2^x$ elements with $x$ an integer such that $8 < x <
+    21$. Thus we have twelve different dataset sizes. We took the system's dictionary file and the
+    list of top level domains from Wikipedia and combined these to generate a list of items in the
+    form \texttt{www.\$word.\$tld}, where \texttt{\$word} is one of words from the the dictionary
+    and \texttt{\$tld} is a top level domain.  Only words consisting of the lowercase letters
+    \texttt{a} through \texttt{z} were used. Constraining the words to consist of 8 to 12 characters
+    and limiting the top-level domains to exactly two letters, all possible combinations were
+    generated.  For the full paths, a number of domains were selected and each was appended a random
+    alphanumeric string of eighty characters.  Finally, for the third type of key, an IP address, we
+    used 32 bit integers. The exact implementation can be viewed in the sourcecode\footnote{ref to
+    apendix or code dump}.
 
@@
 
-    Memory measurements were performed by providing a custom allocator which tracked the number of memory
-    requests.  Note that the latter deviates from our initial plan, where we stated\footnote{ref?}we would
-    use the POSIX API\footnote{ref to the posix api}.  Unfortunately, that turned out to be too
-    coarse-grained for our purposes and we could thus get no accurate measurements.  The
-    allocator-based technique we employed did require some slight modifications to the skip list
-    implementation we used, but this should not have caused a significant slowdown. We assume that any 
-    influence will only have made the tests more fair, as now allocation was done the same way in all 
-    three setups.
+    Memory measurements were performed by providing a custom allocator which tracked the number of
+    memory requests.  Note that the latter deviates from our initial plan, where we
+    stated\footnote{ref?}we would use the POSIX API\footnote{ref to the posix api}.  Unfortunately,
+    that turned out to be too coarse-grained for our purposes and we could thus get no accurate
+    measurements.  The allocator-based technique we employed did require some slight modifications
+    to the skip list implementation we used, but this should not have caused a significant slowdown.
+    We assume that any influence will only have made the tests more fair, as now allocation was done
+    the same way in all three setups.
 
-    After some tests we chose to decrease the number of runs used for measurements.  In the
-    proposal we stated that for each time measurement we would perform the relevant operation 1000
-    times and then take the average of that.  In practice, this made the tests run for longer than
-    we could afford. Because the results given by performing 100 runs were comparable, we settled for that.  
-    We expect the general result to be similar; refer to section\footnote{ref}for details
-    about what \emph{could} have gone wrong.
+    After some tests we chose to decrease the number of runs used for measurements.  In the proposal
+    we stated that for each time measurement we would perform the relevant operation 1000 times and
+    then take the average of that.  In practice, this made the tests run for longer than we could
+    afford. Because the results given by performing 100 runs were comparable, we settled for that.
+    We expect the general result to be similar; refer to section\footnote{ref}for details about what
+    \emph{could} have gone wrong.
 
@@
 
-    In particular, constructing a small hash table\footnote{state the number of elements in a "small" hash table}
-    after constructing many big ones takes a disproportionally long time.
-Only the first hash table is affected, and the time is 			comparable to the time required to construct a hash table with 4000 elements, which is illogical.
-    We assume this is caused by the allocator but the reason for such behaviour is still unclear.  The fact
-    that a hashtable needs contiguous memory is likely to have some influence, as well as the
-    allocator returning memory to the operating system.  However, we reason the latter should also have happened
-    with the other structures.
+    In particular, constructing a small hash table\footnote{state the number of elements in a
+    "small" hash table} after constructing many big ones takes a disproportionally long time.  Only
+    the first hash table is affected, and the time is comparable to the time required to construct a
+    hash table with 4000 elements, which is illogical.  We assume this is caused by the allocator
+    but the reason for such behaviour is still unclear.  The fact that a hashtable needs contiguous
+    memory is likely to have some influence, as well as the allocator returning memory to the
+    operating system.  However, we reason the latter should also have happened with the other
+    structures.
 
-    A similar issue occurs around 4000 or 8000 elements (depending on data structure).  For all three
-    datastructures, the cost of inserting and querying grows at a higher rate than before.  This is difficult to
-    quantify\footnote{Can we do some sort of linear regression to prove this is the case?}, but it
-    appears some extra load is added at that point.  We suspect it is related either to the
-    allocator having to acquire more memory or the CPU cache running out and more cache misses occuring.
-    We would measure this as well, if there was more time.
+    A similar issue occurs around 4000 or 8000 elements (depending on data structure).  For all
+    three datastructures, the cost of inserting and querying grows at a higher rate than before.
+    This is difficult to quantify\footnote{Can we do some sort of linear regression to prove this is
+    the case?}, but it appears some extra load is added at that point.  We suspect it is related
+    either to the allocator having to acquire more memory or the CPU cache running out and more
+    cache misses occuring.  We would measure this as well, if there was more time.
 
@@
     \section{Conclusions}
-    Our main point of research was researching which associative data structure performed best on
-    a various number of insertions, in terms of the speed of insertions and look-ups, and memory usage.
-    In terms of speed we can state with certainty that the hash table is the superior data structure.
-    It outperforms the other two in almost every category, most significantly in larger datastructures where
-    it often takes less than half as much time as the skiplist and less than two thirds the time of the binary search tree.
-    In most cases, the binary search tree is a bit slower, whereas the skip list is the slowest structure in
-    virtually every case. The only motivation for using a hashmap for this problem appears to be in memory usage,
-    although the advantage gained from this is relatively small.
+    Our main point of research was researching which associative data structure performed best on a
+    various number of insertions, in terms of the speed of insertions and look-ups, and memory
+    usage.  In terms of speed we can state with certainty that the hash table is the superior data
+    structure.  It outperforms the other two in almost every category, most significantly in larger
+    datastructures where it often takes less than half as much time as the skiplist and less than
+    two thirds the time of the binary search tree.  In most cases, the binary search tree is a bit
+    slower, whereas the skip list is the slowest structure in virtually every case. The only
+    motivation for using a hashmap for this problem appears to be in memory usage, although the
+    advantage gained from this is relatively small.
 
-    Our secondary point of research encompassed the effect of key complexity on associative data structures.
-    Our different types of keys were IP adresses, domain names and full path URLs. We have seen that as expected,
-    IP adresses as keys provide a significant improvement over the other two, and uses less memory as well. Interesting to
-    note is that we found the only advantage of the binary search tree over the hashmap here, is that insertions of
-    a small amount of IP adresses, up to around $2^13$ insertions, is faster. So when writing an application with very
-    few keys of low complexity, and with more insertions than queries, the binary search tree may be the best option. Finally,
-    we have found that the domain key type does not outperform the full path key type. It is not significantly better
-    if it is better at all.
-    %The skip list uses the least memory but is awful in every other imaginable way.  The hashmap is
-    %usually the best, except when we're dealing with a small number of integer keys and doing far more
-    %inserts than queries.  Note that this doesn't actually happen in practice.
+    Our secondary point of research encompassed the effect of key complexity on associative data
+    structures.  Our different types of keys were IP adresses, domain names and full path URLs. We
+    have seen that as expected, IP adresses as keys provide a significant improvement over the other
+    two, and uses less memory as well. Interesting to note is that we found the only advantage of
+    the binary search tree over the hashmap here, is that insertions of a small amount of IP
+    adresses, up to around $2^13$ insertions, is faster. So when writing an application with very
+    few keys of low complexity, and with more insertions than queries, the binary search tree may be
+    the best option. Finally, we have found that the domain key type does not outperform the full
+    path key type. It is not significantly better if it is better at all.
 
     \section{Reflection}
-    In general, we are content with the results of our research. We were able to provide clear and concise answers to our
-    research questions. Even though our research questions may appear broad, we have defined a good scope in our research
-    proposal and have virtually not altered this scope in our research. One small change is that we also included datasets of
-    size $2^8$ because we saw some interesting results in very small sets. In our implementation we have made some significant
-    changes because we ran into pratical problems. Some of these required us to change the allocator and generate our keys
-    in a slightly different fashion, which we all documented in this paper.
+    In general, we are content with the results of our research. We were able to provide clear and
+    concise answers to our research questions. Even though our research questions may appear broad,
+    we have defined a good scope in our research proposal and have virtually not altered this scope
+    in our research. One small change is that we also included datasets of size $2^8$ because we saw
+    some interesting results in very small sets. In our implementation we have made some significant
+    changes because we ran into pratical problems. Some of these required us to change the allocator
+    and generate our keys in a slightly different fashion, which we all documented in this paper.
 
-    Since we have found that the hash map is the best associative data structure overal for this problem, we could do further
-    research in different kind of hash maps. We might try different hash functions, different methods of collision resolution,
-    and perhaps survey different kinds of concurrent hash maps.
+    Since we have found that the hash map is the best associative data structure overal for this
+    problem, we could do further research in different kind of hash maps. We might try different
+    hash functions, different methods of collision resolution, and perhaps survey different kinds of
+    concurrent hash maps.
 

### 2014-04-17 paper/paper.tex
Reworked conclusion and reflection.

@@
 
-    \subsection*{Data Structures}
+    \subsection{Data Structures}
 
@@
 
-    \subsubsection*{Data}
+    \subsection{Data Sources}
     For our data, we used datasets containing $2^x$ elements with $x$ an integer such that $8 < x <
@@
 
-    \subsection*{Measurements}
+    \subsection{Measurements}
+    \label{subsec:measurements}
     For our time measurements, we used libstdc++'s implementation of the standard
-    \texttt{std::chrono::high\_resolution\_clock} class.
+    \texttt{std::chrono::high\_resolution\_clock} class.  There is no official documentation about
+    the resolution of this clock, but measurements show that it is likely below one microsecond,
+    which is sufficient for our purposes.\footnote{\url{http://stackoverflow.com/a/5524138/559931}}
 
@@
     Our main point of research was researching which associative data structure performed best on a
-    various number of insertions, in terms of the speed of insertions and look-ups, and memory
-    usage.  In terms of speed we can state with certainty that the hash table is the superior data
-    structure.  It outperforms the other two in almost every category, most significantly in larger
-    datastructures where it often takes less than half as much time as the skiplist and less than
-    two thirds the time of the binary search tree.  In most cases, the binary search tree is a bit
-    slower, whereas the skip list is the slowest structure in virtually every case. The only
-    motivation for using a hashmap for this problem appears to be in memory usage, although the
-    advantage gained from this is relatively small.
+    various number of insertions, in terms of the speed of insertions and queries, and in terms of
+    memory necessary per element.  In terms of speed we can state with certainty that the hash table
+    is the superior data structure.  It outperforms the other two in almost every category,
+    particularly in cases where the number of elements is high, where it is roughly two times faster
+    than the skip list and one and a half times faster than the binary search tree.  The binary
+    search tree outperforms the hash table when the number of elements is low, and the number of
+    insertions is relatively larger than the number of queries.  Whether this situation can occur in
+    practice depends on the speed of deletions, which we have not measured; in the given context,
+    however, the conditions are unlikely to be achieved.  A skip list may be preferable if memory
+    usage is more important than query speeds; given that both insert and query operations take less
+    than ten microseconds even in the largest maps we measured, this is likely the case if queries
+    and inserts are rare.
 
     Our secondary point of research encompassed the effect of key complexity on associative data
-    structures.  Our different types of keys were IP adresses, domain names and full path URLs. We
-    have seen that as expected, IP adresses as keys provide a significant improvement over the other
-    two, and uses less memory as well. Interesting to note is that we found the only advantage of
-    the binary search tree over the hashmap here, is that insertions of a small amount of IP
-    adresses, up to around $2^13$ insertions, is faster. So when writing an application with very
-    few keys of low complexity, and with more insertions than queries, the binary search tree may be
-    the best option. Finally, we have found that the domain key type does not outperform the full
-    path key type. It is not significantly better if it is better at all.
+    structures.  Our different types of keys were IP adresses, domain names and full path URLs.  Our
+    measurements confirmed our intuition that IP addresses impose significantly lower costs over the
+    other two key types, as well as requiring less memory.  This improvement was more noticable in
+    the binary search tree than in the two other systems.  The corresponding expected improvement
+    when comparing the domain key type to the full path key type were not observed.
 
@@
     concise answers to our research questions. Even though our research questions may appear broad,
-    we have defined a good scope in our research proposal and have virtually not altered this scope
-    in our research. One small change is that we also included datasets of size $2^8$ because we saw
-    some interesting results in very small sets. In our implementation we have made some significant
-    changes because we ran into pratical problems. Some of these required us to change the allocator
-    and generate our keys in a slightly different fashion, which we all documented in this paper.
+    we have defined a good scope in our research proposal and have not altered the overall goal
+    of our research.  Errors in the implementation lead us to having to discard measurements of 256
+    elements as the outliers were too great to compensate for.  Fortunately, our focus lay primarily
+    on the larger datasets and thus this is not significant for our conclusions.  The implementation
+    we finally arrived at differs from what we outlined, but the changes are largely to increase the
+    accuracy of measurements or the speed at which measurements can be obtained, and have no
+    fundamental influence on the results.  The primary difficulty we encountered was the number of
+    dimensions in our data: with 117 independent measurements, it took considerable time to perform
+    them all, and then to perform statistical analysis on them.
 
-    Since we have found that the hash map is the best associative data structure overal for this
-    problem, we could do further research in different kind of hash maps. We might try different
-    hash functions, different methods of collision resolution, and perhaps survey different kinds of
-    concurrent hash maps.
+    Knowing the strengths and weaknesses of the various data structures, further research could
+    compare subvariants of the hash table and skip list.  In particular, it may be of interest to
+    experiment with differen choices of hash function and collision resolution scheme.  Larger
+    numbers of elements may also provide interesting challenges, as the number of elements may be
+    too great to keep in main memory.  Finally, the situation where a map is accessed by multiple
+    threads concurrently provides extra complications that could be considered.
 

### 2014-04-17 paper/paper.tex
Rewrote plan.

@@
     In this paper, we attempt to answer these questions. Section~\ref{sec:plan} provides details
-    about our research plan and the scopes of our research. Section~\ref{sec:implementation} is
-    about the implementation of our setup in C++.  The rest of the paper states our results and the
+    about the plan and scope of our research. Section~\ref{sec:implementation} is about the
+    implementation of our setup in C++.  The rest of the paper states our results and the
     conclusions that can be drawn from them.
@@
 
-    \subsubsection*{Scope and assumptions}
+    We set out to collect experimental data about three performance characteristics of the data
+    structures we selected: insertion speed, query speed, and memory usage.  We would start by
+    writing code that performed the necessary measurements, run it repeatedly to find the average
+    and standard error, and then perform statistical analysis on the results.
 
-    We make several assumptions for our research.
+    In order to perform the tests we would require a set of keys for every key type.  For IP
+    addresses, consecutive integers starting from 0 would be used.  For domain names and full paths,
+    we wanted to simulate more realistic data.  We would use a word list and a list of top level
+    domains and use possible combinations of elements of each as our domains.  For full paths, we
+    would take some number of domains and append random strings of a fixed length to simulate
+    resource identifiers such as used by YouTube and Google Drive.  In the latter case, we wanted
+    the number of paths per domain to be relatively high, as we expected filtered webpages to be
+    more likely to belong to the same site.  Therefore, instead of taking all possible combinations,
+    we first took a subset of our list of domains and only then added random strings.
 
-    Firstly we plan on only working with datasets that are certain to fit into the main memory of our machine. We intend to
-    purely examine the relative speed and memory-efficiency of the different data structures. Using datasets larger than
-    memory will cause regular lookups to the harddrive, which we assume would skew our results.
+    When running our tests, we would take some subset of the set involved and measure the time it
+    took to run an insert or query using each element of that subset.  Queries and inserts would not
+    be mixed, and each element would be queried and inserted exactly once.  The map to be queried
+    against would contain exactly the elements of the subset.  All in all, this would allow us to
+    compensate for issues that could occur due to particular orders having particular performance
+    characteristics, while not disturbing measurements in ways that a stochastic element count
+    likely would.
 
-    In order to make the measurements more predictable, we will always ensure that insertions are only performed with
-    non-existent keys. The alternative would make either the memory usage or number of insertions non-deterministic, which
-    could cause inaccuracies in measurement. Similarly, we require that all queries be performed on existing keys.
-
-    To mimic real usage, we will insert and then access the keys in an arbitrary order.  However, we will keep the order consistent
-    between tests on different data structures.  Together this will minimize any bias related to any kind of order-dependent
-    performance, which would likely be an advantage specific to the implementation, not to the data structure. We will
-    generate the order for both lookup and insertion by shuffling the dataset we chose for the specific test. This means that
-    orders will not be persistent between different data set sizes; however, we do not expect this to be significant.
-
-    We realize that methods to measure the amount of time required for actions will themselves need time. We will however
-    use the same time-measuring method for all data structures, so we can still use the results to compare the different
-    data structures.
+    We would ensure that all subset fit in main memory, as the alternative would significantly
+    favour data structures that used less memory.  Furthermore, we would perform all time
+    measurements repeatedly and then take the average in order to decrease the influence of timing
+    overhead.
 
@@
 
-    We used C++ implementations of the data structures involved, and also the language for our
+    We used C++ implementations of the data structures involved, and also used the language for our
     measuring code.  We chose C++ as it is an industry standard\footnote{ref}, provides the

### 2014-04-17 paper/paper.tex
Results section

--- 2014-04-17 paper/paper.tex
@@
 \usepackage{hyperref}
+\usepackage[pdftex]{graphicx}
 
@@
 
-    Our results were great!  They are best analysed with a two-sample independent T test.  We had an
-    equal number of measurements about group 1 (with mean $x_1$ and standard deviation $\sigma_1$ and
-    about group 2 (with mean $x_2$ and standard deviation $\sigma_2$).  From this we can conclude that
-    data structure 1 is, on average, unquestionably superior.
+    Now that we have done the tests we have gained a lot of data that we can analyze and base our 
+    conclusions on. To do this, we will need to make our data more easily readable and use statistical
+    tests to draw conclusions.
+
+     First, to get an overview of our data, we will determine the averages for each scenario
+    (Datastructure, keytype and number of entries). For these averages we will look at operation speed per
+    element and memory usage per element. These can be seen in appendix \ref{App:AveResults}.
+
+    Now we may notice patterns in these averages, but to actually draw conclusions from the results, we will
+    need to use statistical methods. Depending on what results we want to compare there are different statistical
+    tests we can use. For our results we shall make use of independent samples T test and bivariate correlation test.
+
+    We shall use a significance of $0.95$ and we perform all statistical tests using the SPSS software package.
+
+    \subsection{Correlation between number of entries and results}
+    First we look at the influence the number of entries has on the per element speed of operations. For this
+    we will use a bivariate correlation test, since we want to check if there exists a correlation between 
+    the number of entries and the speed of operations.
+
+    Using analysis in SPSS it turns out that for almost all situations there exists positive correlations between the number
+    of entries and the speed of both queries and inserts. For both skiplist and BST these correlations are significant for all
+    keytypes. In all cases they have positive correlations between $0.4$ and $0.6$.
+
+    However in the case of hashmaps we don't always have correlations. For all situations there does exist a positive correlation
+    between the number of entries and the per element query time. But for inserts there only exists a correlation when we 
+    look at Ip-address keys. If we look at Domain or Full path keys, there is no significant correlation between the number of entries
+    and the per element speed of insert operation.
+
+    \subsection{Difference in performance between different keytypes}
+    Another part of our research question was: ``How does the complexity of the key influence the performance of 
+    various associative data structures?". Thus we would like to compare performance between the different keytypes. 
+    To look at this, we will set our number of entries to 4096 and compare the different key types.
+
+    It turns out the results are largely the same for the different data structures. For all of the data structures the use of
+    Ip-addresses is significantly faster in both query and insert speeds and has lower memory usage as well, this is even true
+    no matter the number of entries.
+
+    For the difference between full path and domain the results become more complex. Looking at BST we get the following results:
+    Insert speed on full paths is significantly faster than insert speed on domains, however query speed on domains is significantly faster
+    than query speed on full paths.
+
+    Looking at skiplist we get the following results: Insert speed on full paths is significantly faster than insert speed on domains. Query
+    speed however does not significantly differ between domains and full paths. If we look at higher number of entries query speed becomes 
+    significantly faster for full paths.
+
+    Finally for our results with hashmap: Domains are significantly faster in both insert and query speed compared to full paths. 
+    This is true no matter the number of entries.
+
+    \subsection{Comparison between different data structures}
+    We would like to be able to use our results to determine which data structures are better in which cases. 
+    For this we will have to compare the different data structures for specific situations. We will use an independent
+    samples T test for this, since our data points are independent of each other.
+
+    First off, for almost all situations with the keytypes Full path and Domain, hashmap is significantly faster than BST and skiplist
+    in both query speed and in insert speed. The only exception is in the case of 256 entries, which is most likely a
+    problem in our setup.
+
+    More interesting are the speeds in the case of the Ip-addresses. While the hashmap is still significantly faster 
+    for query speeds, no matter the number of entries, there is a different result for the insert speeds. For 4096
+    entries or less, the BST significantly outperforms the hashmap in terms of insert speed. For any number of 
+    entries higher than that, the hashmap outperforms the BST again.
 
@@
     slower, whereas the skip list is the slowest structure in virtually every case. The only
-    motivation for using a hashmap for this problem appears to be in memory usage, although the
+    motivation for using a skiplist for this problem appears to be in memory usage, although the
     advantage gained from this is relatively small.
@@
 
+    \newpage
+    \appendix
+    \section{Average results} \label{App:AveResults}
+
+    \begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth]{Ip-address_averages.png}
+    \caption{The average per element results for each situation with Ip-adresses.}
+    \end{figure}
+
+    \begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth]{Domain_averages.png}
+    \caption{The average per element results for each situation with domain adresses.}
+    \end{figure}
+
+    \begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth]{Full_path_averages.png}
+    \caption{The average per element results for each situation with full path adresses.}
+    \end{figure}
+
 \end{document}

### 2014-04-17 paper/paper.tex
Rewrote data sources and measurements.

--- 2014-04-17 paper/paper.tex
@@
     their respective permissions, be it whitelisted or blacklisted for certain users.  We have
-    chosen to compare the performance of the binary search tree, the hash map, and the skip list, as
+    chosen to compare the performance of the binary search tree, the hash table, and the skip list, as
     these are commonly employed in practice.  It is of particular interest how each data structure
@@
 
+    For memory measurements, such repeated runs were deemed unnecessary as the deviation there was
+    likely to be lower; a single map was thus constructed, and the difference in memory usage as
+    shown by the POSIX API would be recorded.  As discussed in Subsection~\ref{subsec:measurements},
+    this turned out to be unfeasible.
+
     We would ensure that all subset fit in main memory, as the alternative would significantly
@@
 
+    The following is an overview of our implementation.  More specific details can be found in the
+    source code.\footnote{cite code here}
+
     \subsection{Data Structures}
 
-    Using C++ as the implementation language, we chose libstdc++'s implementations of the standard class
-    templates \texttt{std::map} and \texttt{std::unordered\_map} for the binary search
-    tree and the hash table respectively.  As the C++ standard library does not contain a skip
+    With C++ as the implementation language, we chose libstdc++'s implementations of the standard
+    class templates \texttt{std::map} and \texttt{std::unordered\_map} for the binary search tree
+    and the hash table respectively.  As the C++ standard library does not contain a skip
     list\footnote{reference to the standard here}, we used an open-source third party library,
@@
     \subsection{Data Sources}
-    For our data, we used datasets containing $2^x$ elements with $x$ an integer such that $8 < x <
-    21$. Thus we have twelve different dataset sizes. We took the system's dictionary file and the
-    list of top level domains from Wikipedia and combined these to generate a list of items in the
-    form \texttt{www.\$word.\$tld}, where \texttt{\$word} is one of words from the the dictionary
-    and \texttt{\$tld} is a top level domain.  Only words consisting of the lowercase letters
-    \texttt{a} through \texttt{z} were used. Constraining the words to consist of 8 to 12 characters
-    and limiting the top-level domains to exactly two letters, all possible combinations were
-    generated.  For the full paths, a number of domains were selected and each was appended a random
-    alphanumeric string of eighty characters.  Finally, for the third type of key, an IP address, we
-    used 32 bit integers. The exact implementation can be viewed in the sourcecode\footnote{ref to
-    apendix or code dump}.
+
+    We performed tests using subsets of twelve different sizes, all of the form $2^x$ and with $8 <
+    x \le 20$.  The word list was obtained by taking the system's dictionary file and selecting all
+    words that contained only lowercase letters and that were at least 8 and at most 12 letters
+    long.  The top-level domain list was obtained from Wikipedia, and filtered down to only those
+    domains that consisted of two letters from the English alphabet.  The resulting domains were of
+    the form \texttt{www.\$word.\$tld}, where \texttt{\$word} is one of words from the the word list
+    and \texttt{\$tld} is a top level domain.  For the full paths, a number of domains were selected
+    and each was appended a random alphanumeric string of eighty characters.  By using the same
+    domain with multiple random strings, a sufficiently large number of paths was generated.
+    Finally, as described in the research plan, consecutive 32-bit integers starting from zero were used
+    for the IP addresses.
 
@@
     \label{subsec:measurements}
+
     For our time measurements, we used libstdc++'s implementation of the standard
@@
     Memory measurements were performed by providing a custom allocator which tracked the number of
-    memory requests.  Note that the latter deviates from our initial plan, where we
-    stated\footnote{ref?}we would use the POSIX API\footnote{ref to the posix api}.  Unfortunately,
-    that turned out to be too coarse-grained for our purposes and we could thus get no accurate
-    measurements.  The allocator-based technique we employed did require some slight modifications
-    to the skip list implementation we used, but this should not have caused a significant slowdown.
-    We assume that any influence will only have made the tests more fair, as now allocation was done
-    the same way in all three setups.
+    memory requests.  Note that the latter deviates from our initial plan, where we stated we would
+    use the POSIX API.  Unfortunately, the measurements provided by that turned out to be too
+    coarse-grained for our purposes, and our initial measurements were unusable.  The
+    allocator-based technique we employed did require some slight modifications to the skip list
+    implementation we used, but this should not have caused a significant slowdown.  We assume that
+    any influence will only have made the tests more fair, as now allocation was done the same way
+    in all three structures.
 
-    After some tests we chose to decrease the number of runs used for measurements.  In the proposal
-    we stated that for each time measurement we would perform the relevant operation 1000 times and
-    then take the average of that.  In practice, this made the tests run for longer than we could
-    afford. Because the results given by performing 100 runs were comparable, we settled for that.
-    We expect the general result to be similar; refer to section\footnote{ref}for details about what
-    \emph{could} have gone wrong.
+    After running a number of tests we chose to decrease the number of times we repeated each run.
+    We had set out to perform each measurement 1000 times and take the average over that.  In
+    parctice, however, the results obtained by performing only 100 runs did not vary greatly, and
+    allowed us to perform more individual measurements.
 

### 2014-04-17 paper/paper.tex
Cleaned up filenames.

@@
 
+    \newpage
+    \appendix
+    \section{Average results} \label{App:AveResults}
+
+    \begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth]{ip-address_averages.png}
+    \caption{The average per element results for each situation with Ip-adresses.}
+    \end{figure}
+
+    \begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth]{domain-averages.png}
+    \caption{The average per element results for each situation with domain adresses.}
+    \end{figure}
+
+    \begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth]{full-path-averages.png}
+    \caption{The average per element results for each situation with full path adresses.}
+    \end{figure}
+
 \end{document}

### 2014-04-17 paper/paper.tex
Fixed up results.

--- 2014-04-17 paper/paper.tex
@@
 
-    We performed tests using subsets of twelve different sizes, all of the form $2^x$ and with $8 <
-    x \le 20$.  The word list was obtained by taking the system's dictionary file and selecting all
-    words that contained only lowercase letters and that were at least 8 and at most 12 letters
+    We performed tests using subsets of thirteen different sizes, all of the form $2^x$ and with $8
+    \le x \le 20$.  The word list was obtained by taking the system's dictionary file and selecting
+    all words that contained only lowercase letters and that were at least 8 and at most 12 letters
     long.  The top-level domain list was obtained from Wikipedia, and filtered down to only those
@@
     domain with multiple random strings, a sufficiently large number of paths was generated.
-    Finally, as described in the research plan, consecutive 32-bit integers starting from zero were used
-    for the IP addresses.
+    Finally, as described in the research plan, consecutive 32-bit integers starting from zero were
+    used for the IP addresses.
 
@@
 
-    Our results were great!  They are best analysed with a two-sample independent T test.  We had an
-    equal number of measurements about group 1 (with mean $x_1$ and standard deviation $\sigma_1$ and
-    about group 2 (with mean $x_2$ and standard deviation $\sigma_2$).  From this we can conclude that
-    data structure 1 is, on average, unquestionably superior.
+    Having done the tests, we have obtained a lot of data that we can base our conclusions on.  A
+    preliminary overview can be obtained in Appendix~\ref{App:AveResults}, which lists the average
+    per-element results we obtained.  However, in order to draw conclusions from our data we will
+    require statistical methods.  For our results, we shall make use of the independent samples
+    T-test and the bivariate correlation test.  In all cases, we will use a significance level of
+    $0.05$ and perform the tests using the SPSS software package.
+
+    Note that in the following analyses, we do not consider the results of the $2^8$ group, as the
+    experimental errors involved were too significant.
+
+    \subsection{Correlation between number of entries and results}
+
+    Firstly, we shall look at the influence the number of entries has on the per element speed of
+    operations.  For this we will use a bivariate correlation test, since we want to check if there
+    exists a correlation between two variables of ratio levels of measurements.
+
+    Using the analysis functions of SPSS shows that for almost all situations there exists a
+    positive correlation between the number of entries and the duration of both queries and inserts.
+    For both the skip list and the binary search tree, these correlations are significant for all
+    key types. In all cases they have positive correlation coefficients between $0.4$ and $0.6$.
+
+    In the case of hash table, however, we do not always find a correlation.  For all situations there does
+    exist a positive correlation between the number of entries and the per element query duration, but
+    the insertion duration is only correlated to the number of entries in the case of the IP address
+    key type.  For the domain and full path key types, no correlation between the number of entries
+    and the duration of insertion operations is observable.
+
+    \subsection{Difference in performance between different keytypes}
+
+    In order to answer the second part of our research question, we must compare the behaviour of
+    the data structures given different key types.  We start by fixing the number of entries; we
+    chose to do this at 4096 elements.  We then again performed an independent samples T-test
+    using SPSS.
+
+    The results are largely the same for all three data structures. For all of the data structures
+    the use of IP addresses gives significantly higher speeds both for query and insertion, and the
+    memory usage is lower.  These results can be replicated for all entry counts.
+
+    The difference between the results for the full path and the domain key types is less clear-cut.
+    In the case of the binary search tree, the insertion speed for full paths is significantly higher
+    than the insertion speed for domains, while the opposite is true for query speeds.  The skip
+    list also has a significantly higher insertion speed for full paths than for domains, but there
+    is no significant difference for query speed unless we increase the number of elements.
+
+    The hash table displays the expected behaviour, and both the insertion and query speeds for
+    domains are significantly higher than for full paths.
+
+    \subsection{Comparison between different data structures}
+
+    We would like to be able to use our results to determine which data structures are better in
+    which cases.  For this we will have to compare the different data structures for specific
+    situations. We will use an independent samples T test again for this, since our data points are
+    independent of each other.
+
+    For all element counts and domain or full path key types, the hash table has significantly higher insertion
+    and query speeds than the binary search tree and skip list.  For IP addresses, the results are
+    more interesting.  While the hash table is still significantly faster for query speeds for any
+    number of entries, the binary search tree wins on insertion speed when there are 4096 entries or
+    less.
 
@@
 
-    In particular, constructing a small hash table\footnote{state the number of elements in a
-    "small" hash table} after constructing many big ones takes a disproportionally long time.  Only
-    the first hash table is affected, and the time is comparable to the time required to construct a
-    hash table with 4000 elements, which is illogical.  We assume this is caused by the allocator
-    but the reason for such behaviour is still unclear.  The fact that a hashtable needs contiguous
-    memory is likely to have some influence, as well as the allocator returning memory to the
-    operating system.  However, we reason the latter should also have happened with the other
-    structures.
+    In particular, constructing a small hash table after constructing many big ones reliably took a
+    disproportionally long time.  Only the first hash table is affected, and the time is comparable
+    to the time required to construct a hash table with 4000 elements.  We assume this is caused by
+    the allocator but the reason for such behaviour is still unclear.  The fact that a hash table
+    needs contiguous memory is likely to have some influence, as well as the allocator returning
+    memory to the operating system.  However, we reason the latter should also have happened with
+    the other structures.
 
@@
     three datastructures, the cost of inserting and querying grows at a higher rate than before.
-    This is difficult to quantify\footnote{Can we do some sort of linear regression to prove this is
-    the case?}, but it appears some extra load is added at that point.  We suspect it is related
+    This is difficult to quantify, but it appears some extra load is added at that point.  We suspect it is related
     either to the allocator having to acquire more memory or the CPU cache running out and more
-    cache misses occuring.  We would measure this as well, if there was more time.
+    cache misses occuring.
 
-    A potentially related issue is that the standard deviation of our measurements rises around the
-    same point.  That increases the likelihood of it being allocator-related, as that involves a
-    system call and could cause scheduling issues\footnote{elaborate/ref?}, despite the system being
-    otherwise very close to idle.
-
-    Also, we acquired counterintuitive results showing that inserting or querying a full path is
-    faster than querying only the domain name.  This can partially be explained\footnote{Not really,
-    but ehhh.}by the fact that the string implementation we used was based on reference counting.
-    %(If it'd also provide SSO, this would be easy to explain, but the darn thing doesn't.)
+    This is also a possible explanation that the results we obtained which claimed that operations
+    for a full path key are more efficient than identical operations for a domain key.  An
+    alternative explanation is that the differences in the data sets used for the keys made
+    the comparison cheaper for the full paths, despite their greater size.  However, the way we
+    constructed the sets would suggest the opposite.
 
@@
 
+    \newpage
+    \appendix
+    \section{Average Results} \label{App:AveResults}
+
+    \begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth]{ip-address-averages.png}
+    \caption{Per element averages with the IP-address key type.}
+    \end{figure}
+
+    \begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth]{domain-averages.png}
+    \caption{Per element averages with the domain key type.}
+    \end{figure}
+
+    \begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth]{full-path-averages.png}
+    \caption{Per element averages with the full path key type.}
+    \end{figure}
+
 \end{document}

### 2014-04-17 paper/paper.tex
typos

@@
         We have researched the performance of the hash table, binary search tree, and skip list
-        associative datastructures in the context of a web filter.  Our measurements were made with
+        associative data structures in the context of a web filter.  Our measurements were made with
         structures of various sizes (ranging from $2^8$ up to $2^{20}$) and with different possible
@@
     sites a user may not visit unless they have certain credentials, or alternatively, keeping a
-    whitelist and prohibiting any sites not on it.  As the world-wide web is growing, these lists
+    whitelist and prohibiting any sites not on it.  As the world wide web is growing, these lists
     have grown larger to a point where the speed of the underlying data structure has become a key
@@
 
-    While the decision of data structure might mostly be dependant on the number of entries, the
+    While the decision of data structure might mostly be dependent on the number of entries, the
     complexity of the key might also be of concern. Where permissions per page would be preferred, a
-    more resource optimal solution might be found in filtering only on domain or even IP adress.  In
+    more resource optimal solution might be found in filtering only on domain or even IP address.  In
     researching the influence of this we formulate a secondary research question:
@@
     \begin{quotation}
-        How does the complexity of the key influence the preformance of various associative data structures?
+        How does the complexity of the key influence the performance of various associative data structures?
     \end{quotation}
@@
     We had set out to perform each measurement 1000 times and take the average over that.  In
-    parctice, however, the results obtained by performing only 100 runs did not vary greatly, and
+    practice, however, the results obtained by performing only 100 runs did not vary greatly, and
     allowed us to perform more individual measurements.
@@
 
-    \subsection{Difference in performance between different keytypes}
+    \subsection{Difference in performance between different key types}
 
@@
     A similar issue occurs around 4000 or 8000 elements (depending on data structure).  For all
-    three datastructures, the cost of inserting and querying grows at a higher rate than before.
+    three data structures, the cost of inserting and querying grows at a higher rate than before.
     This is difficult to quantify, but it appears some extra load is added at that point.  We suspect it is related
     either to the allocator having to acquire more memory or the CPU cache running out and more
-    cache misses occuring.
+    cache misses occurring.
 
@@
     Our secondary point of research encompassed the effect of key complexity on associative data
-    structures.  Our different types of keys were IP adresses, domain names and full path URLs.  Our
+    structures.  Our different types of keys were IP addresses, domain names and full path URLs.  Our
     measurements confirmed our intuition that IP addresses impose significantly lower costs over the
-    other two key types, as well as requiring less memory.  This improvement was more noticable in
+    other two key types, as well as requiring less memory.  This improvement was more noticeable in
     the binary search tree than in the two other systems.  The corresponding expected improvement
@@
     compare subvariants of the hash table and skip list.  In particular, it may be of interest to
-    experiment with differen choices of hash function and collision resolution scheme.  Larger
+    experiment with different choices of hash function and collision resolution scheme.  Larger
     numbers of elements may also provide interesting challenges, as the number of elements may be

### 2014-04-17 paper/paper.tex
Added references.

--- 2014-04-17 paper/paper.tex
@@
 \usepackage{hyperref}
+\usepackage{url}
 \usepackage[pdftex]{graphicx}
@@
     their respective permissions, be it whitelisted or blacklisted for certain users.  We have
-    chosen to compare the performance of the binary search tree, the hash table, and the skip list, as
-    these are commonly employed in practice.  It is of particular interest how each data structure
-    will scale as the number of elements rises.  We have thus chosen for our main research question
-    the following:
+    chosen to compare the performance of the binary search tree, the hash table, and the skip list
+    data structures~\cite{Algorithms} as these are commonly employed in practice.  It is of
+    particular interest how each data structure will scale as the number of elements rises.  We have
+    thus chosen for our main research question the following:
     % Guys, did you notice we did absolutely nothing about predicting this?
@@
     We used C++ implementations of the data structures involved, and also used the language for our
-    measuring code.  We chose C++ as it is an industry standard\footnote{ref}, provides the
+    measuring code.  We chose C++ as it is an industry standard~\cite{TC++PL}, provides the
     necessary tools, and does not carry the risk of garbage collection pauses or effects due to
@@
     The following is an overview of our implementation.  More specific details can be found in the
-    source code.\footnote{cite code here}
+    source code~\cite{OurCode}.
 
@@
 
-    With C++ as the implementation language, we chose libstdc++'s implementations of the standard
+    With C++ as the implementation language, we chose \texttt{libstdc++}'s implementations of the standard
     class templates \texttt{std::map} and \texttt{std::unordered\_map} for the binary search tree
-    and the hash table respectively.  As the C++ standard library does not contain a skip
-    list\footnote{reference to the standard here}, we used an open-source third party library,
-    \texttt{CSSkipList}\footnote{reference to library home page here}.  We also used libstdc++'s
-    implementation of \texttt{std::string} for our string handling\footnote{specify?}needs.
+    and the hash table respectively~\cite{C++11}.  As the C++ standard library does not contain a
+    skip list, we used an open-source third party library, \texttt{CSSkipList}~\cite{CSSkipList}.
+    We also used \texttt{libstdc++}'s implementation of \texttt{std::string} for the storage of
+    domains and full paths.
 

### 2014-04-17 paper/paper.tex
Got rid of a footnote.

@@
     the resolution of this clock, but measurements show that it is likely below one microsecond,
-    which is sufficient for our purposes.\footnote{\url{http://stackoverflow.com/a/5524138/559931}}
+    which is sufficient for our purposes.
 
